---
title: "Study Bounds"
author: "Jill Deines"
date: "June 29, 2017"
output: 
  html_document:
    toc: yes
---

Goal: Document full production of study area boundary, and evaluate the number of counties various options include for a robust county validation. This aggregates study bound polygon generation from several sources prior to manuscript revisions. It also incorporates the area calculations from the old 0.01_polygonAreaFinder.R script.

```{r knitrOpts, echo=FALSE}
library(knitr)
opts_chunk$set(cache=TRUE,
               fig.path='figure/0.02_studyBounds/')
```

**R Packages Needed**

```{r packages, message=FALSE, echo=TRUE}
library(rgdal)
library(rgeos)
library(raster)
```


## Model boundary: uniform buffer
This is a new boundary created 6/29/2017 for manuscript revisions. Load RRCA groundwater model boundary and RRB basin boundary, union, then apply a uniform buffer to the full region.

Buffer distances will be calculated using `CRS('+init=epsg:5070'))`, CONUS albers equal area, to be consistent with final map outputs.

```{r loadBounds}
gisDir <- 'S:/Users/deinesji/HPA/gis'
# load bound used for first submission for comparison
original <-  readOGR(gisDir, 'BigExtent_RRB_RRCA_buff', verbose=F)

# rrca bound exported from model
bound <- readOGR(paste0(gisDir,'/RRCA_Model_Export'), 'bound', verbose=F)

# republican basin
rrb <- readOGR(gisDir, 'RepBasin', verbose=F)

# reproject
aeaProj <- '+init=epsg:5070'
bound <- spTransform(bound, CRS(aeaProj)) 
rrb <- spTransform(rrb, CRS(aeaProj))
original <- spTransform(original, CRS(aeaProj))

# union
unioned <- gUnion(bound,rrb)

# buffer options --------------------------------------------------
# get lateral width
boundwidth <- abs(bbox(unioned)[2,1]-bbox(unioned)[2,2])
percent10 <- boundwidth*.1

# buffer and reformat
buffer10p <- gBuffer(unioned, width=percent10)
buffer10p <- SpatialPolygonsDataFrame(buffer10p, match.ID=F,
                                         data.frame(name='buffered'))

# minimum buffer: rrb matching
buff <- 10000
bufferSmall <- gBuffer(unioned, width=buff)
bufferSmall <- SpatialPolygonsDataFrame(bufferSmall, match.ID=F,
                                         data.frame(name='buffered'))
# make minute adjustments since original data was buffered in HPA class projection
bufferSmall2 <- raster::intersect(bufferSmall, original)

# visualize
par(mar=c(2,0,0,0))
plot(original, border='blue')
plot(buffer10p, border = 'red',add=T)
plot(bufferSmall2, add=T)
plot(unioned, col='gray80',border='gray80',add=T)
legend('bottom',legend = c('original','10 percent','10 km','union'), 
       col = c('blue','red','black','gray80'), lty=1, ncol=4)
#plot(unioned, add=T)
```

## County Analyses
See how buffer choices affect the number of fully included counties

```{r countyCheck}
# county shapefile
tigerDir <- 'S:/Data/GIS_Data/Downloaded/TIGER/TIGER2012/COUNTY'
tigerName <- 'tl_2012_us_county'
counties <- readOGR(tigerDir, tigerName, verbose=F)
counties <- spTransform(counties, CRS(aeaProj))

# subset columns and add 5 digit fips
columnsToKeep <- c('STATEFP','COUNTYFP','COUNTYNS','NAME')
counties <- counties[,columnsToKeep]
counties$fips5 <- paste0(counties$STATEFP, counties$COUNTYFP)


# function to return counties 100% within polygons
getInCounties <- function(aoi0, countyAll, threshold = .999, tolerance = -100){
  #do a slight negative buffer to avoid line mismatches
  aoi <- gBuffer(aoi0, width=tolerance)
  # subset counties by aoi (all counties with any overlap)
  counties.overlap <- countyAll[aoi,]
  # crop counties by aoi
  countyCrop  <- raster::intersect(countyAll, aoi)
  # find the proportion of each county within the aoi by getting the area of
  # both and dividing
  counties.overlap$area_km2 <- round(gArea(counties.overlap, byid=T)/1000000)
  countyCrop$area_km2 <- round(gArea(countyCrop, byid=T)/1000000)
  countyCrop$proportion <- countyCrop$area_km2/counties.overlap$area_km2
  countiesIn <- countyCrop[countyCrop$proportion > threshold,]
  return(countiesIn)
}

# get counties by boundary: original
county.original <- getInCounties(original, counties, threshold=.999)
nrow(county.original)
plot(counties[original,], main = 'Original 35 Counties Used')
plot(original, border = 'blue', add=T)
plot(county.original, col='red', add=T)

# get counties by boundary: strict untion
county.union <- getInCounties(unioned, counties, threshold=.999)
nrow(county.union)
plot(counties[unioned,], main = 'Unioned Counties: 21')
plot(unioned, border = 'blue', add=T)
plot(county.union, col='red', add=T)

# get counties by boundary: smallest consensus buffer
county.10km <- getInCounties(bufferSmall2, counties, threshold=.999)
nrow(county.10km)
plot(counties[bufferSmall2,], main = '10 km buffer Counties: 25')
plot(bufferSmall2, border = 'blue', add=T)
plot(county.10km, col='red', add=T)

# get counties by boundary: biggest consensus buffer
county.10p <- getInCounties(buffer10p, counties, threshold=.999)
nrow(county.10p)
plot(counties[buffer10p,], main = '10% buffer Counties: 34')
plot(buffer10p, border = 'blue', add=T)
plot(county.10p, col='red', add=T)

# counties combined
par(mar=c(2,0,0,0))
plot(county.original, border='green', main = 'County Comparison')
plot(county.10p, col = 'gray20', add=T)
plot(county.original, border='green', add=T)
plot(county.10km, col='blue', add=T)
plot(county.union, col='red', add=T)
legend('bottom', legend = c('original','10%','10 km','union'),
       fill = c('green','gray20','blue','red'), ncol=4)
```

## Effect of choices

