---
title: "Select Counties for Validation Points"
author: "Jill Deines"
date: "Saturday, October 29, 2016"
output: 
  html_document:
    toc: yes
---

Goal: Use Koeppengeiger and USDA Ecoregion climate divisions to sample counties for validation points

Update 11/13/2016: No longer sampling points on a county basis; see 1.15_makeValidationPointsRRB.Rmd for sampling across climate zones and crops.


```{r knitrOpts, echo=FALSE}
library(knitr)
opts_chunk$set(cache=TRUE, cache.path='cache/1.05_sampleCounty/',
               fig.path='figure/1.05_sampleCounty/')
```

**R Packages Needed**

```{r packages, message=FALSE, echo=TRUE}
library(rgdal)
library(raster)
library(dplyr)
```

## Load data

```{r loadData}
# filepaths -----------------------------------------------------
# county shapefile
tigerDir <- 'S:/Data/GIS_Data/Downloaded/TIGER/TIGER2012/COUNTY'
tigerName <- 'tl_2012_us_county'

# aquifer and basin boundaries
gisDir <- 'S:/Users/deinesji/HPA/gis'
rrbName <- 'BigExtent_RRB_RRCA_buff'

# usda ecogreions
ecoDir <- 'S:/Users/deinesji/HPA/data/climate/climateDivisions/USDA_Ecoregions'
ecoName <- 'eco_us'

# koeppen
koepPath <- 'S:/Users/deinesji/HPA/data/climate/climateDivisions/Koeppengeiger/KoeppenGeiger_UScounty.txt'

# read in
counties <- readOGR(tigerDir, tigerName, verbose=F)
rrb <- readOGR(gisDir, rrbName, verbose=F)
ecoregions <- readOGR(ecoDir, ecoName, verbose=F)
koepdata <- read.delim(koepPath)
```


### Get Counties
Clips the TIGER 2012 county shapefile to area of interest and adds the counties Koeppengeiger code

```{r getCounties}
# reproject to wgs84
rrb.ll <- spTransform(rrb, CRS(proj4string(counties)))

# clip counties to the rrb
countyRRB <- intersect(counties,rrb.ll)

# trim dataframe and create 5-digit FIPS
countysub <- countyRRB[,c('STATEFP','COUNTYFP','NAME')]
countysub$STATEFP <- as.character(countysub$STATEFP)
countysub$COUNTYFP <- as.character(countysub$COUNTYFP)
countysub$fips <- paste0(countysub$STATEFP, countysub$COUNTYFP)

# add the koeppengeiger code
koepdata$FIPS <- as.character(koepdata$FIPS)
countyClim <- merge(countysub, koepdata, by.x = 'fips', by.y = 'FIPS', all.x = T)
```




