---
title: "Create Validation Points"
author: "Jill Deines"
date: "Thursday, May 19, 2016"
output: 
  html_document:
    toc: yes
---

Goal: Create a set of validation points to be used in evaluating irrigation maps produced from remote sensing classifications. The aim is to have random points stratified among different crop types as well as non-ag land.

```{r knitrOpts, echo=FALSE}
library(knitr)
opts_chunk$set(cache=TRUE, cache.path='cache/1.1_vPOints/',
               fig.path='figure/1.1_vPoints/')
```

**R Packages Needed**

```{r packages, message=FALSE, echo=TRUE}
library(rgdal)
library(raster)
```

## CDL points
Use the 2011 CDL to stratify random points among crops of interest

Dominant crops:

* Corn
* Winter wheat
* Soybeans
* Sorghum
* Alfalfa
* Millet
* Grass/Pasture
* Fallow/Idle Cropland

```{r loadStuff}
#2011 cdl and key
cdl <- raster('S:/Users/deinesji/HPA/gis/CDL/CDL_RRB_2011_adj.tif')
cdlKey <- read.csv('S:/Users/deinesji/HPA/gis/CDL/CDL_key_2014.csv')

# projection datum not read in, fix it
tweakProj <- "+proj=aea +lat_1=29.5 +lat_2=45.5 +lat_0=23 +lon_0=-96 +x_0=0 +y_0=0 +ellps=GRS80 +datum=NAD83 +units=m +no_defs +towgs84=0,0,0"
proj4string(cdl) <- tweakProj


```

