---
title: "Random Forest"
author: "Jill Deines"
date: "January 17, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, cache=TRUE, cache.path='cache/1.18_rrbRandFor/',
               fig.path='figure/1.18_rrbRandFor/')
```

Goal: Run a variable importance test on a random forest of my training points

randomForest package implementation:
 reference implementation based on CART trees
(Breiman, 2001; Liaw and Wiener, 2008)
– for variables of different types: biased in favor of
continuous variables and variables with many categories
(Strobl, Boulesteix, Zeileis, and Hothorn, 2007)


**R Packages Needed**

```{r packages, message=FALSE, echo=TRUE}
library(randomForest)
library(varSelRF)
```

## Load Data
Load training point data used in GEE classification. These were processed in 1.16_trainingPoints_RRB.Rmd, which involved combining 2010 and 2012 points and removing extraneous columns, so that only the variables and a masterType and masterNum columns remained. I wrote out a csv version of the final kml to be used here.

```{r loadData}
# load kml of combined 2010/2012 training points
trainingFolder <- 'C:/Users/deinesji/Dropbox/1PhdJill/hpa/irrigation/data/GIS/training'
trainingName <- '2010_2012_COMBINED_training_12class_v2.csv'

points <- read.csv(paste0(trainingFolder, '/', trainingName))

# re-class the master types into 3 categories: irrigated, dryland crop, noncrop
typeConverter <- data.frame(masterType = unique(points$masterType),
                            classes = c('irrigated','dryland','irrigated','dryland',
                                        'irrigated','dryland','irrigated','dryland',
                                        'irrigated','dryland','noncrop','noncrop'))
points2  <- merge(points, typeConverter)

# remove extra columns
points3 <- points2[ , -which(names(points2) %in% c('masterType','masterNum'))]

# make a dataset without dryland soy
pointsNoSoy <- points2[points2$masterType != 'soy_dryland',]
pointsNoSoy2 <- pointsNoSoy[ , -which(names(pointsNoSoy) %in%
                                        c('masterType','masterNum'))]

# make an annual variable dataset without dryland soy
columnsWanted <- c('EVI_max_14','EVI_range_14','GI_max_14','GI_range_14',
                   'NDVI_max_14','NDVI_range_14','NDWI_max_14','NDWI_range_14',
                   'aridity','b1','b1_1','greenArid','ndwi_gi','pdsi_ann',
                   'pdsi_grow','pr_ann','pr_early','pr_grow','pr_paw','slope_mean',
                   'classes')
annualNoSoy <- pointsNoSoy2[, columnsWanted]
```

## Random Forest
First I do a random forest that mimics my GEE test (500 trees)
```{r rf500, fig.height = 8}
# run a random forest on all 57 variables
set.seed(415)
fit500 <- randomForest(x = points3[,1:57], # all variable columns (excludes classes)
                    y = as.factor(points2$classes),
                    ntree = 500,
                    importance = TRUE)

# look at variable importance
varImpPlot(fit500)
```

That was fast, let's do more trees

```{r rf1000, fig.height = 8}
# run a random forest on all 57 variables
set.seed(415)
fit1000 <- randomForest(x = points3[,1:57], # all variable columns (excludes classes)
                    y = as.factor(points2$classes),
                    ntree = 1000,
                    importance = TRUE)

# look at variable importance
varImpPlot(fit1000)
```

```{r rf2000, fig.height = 8}
# run a random forest on all 57 variables
set.seed(415)
fit2000 <- randomForest(x = points3[,1:57], # all variable columns (excludes classes)
                    y = as.factor(points2$classes),
                    ntree = 2000,
                    importance = TRUE)

# look at variable importance
varImpPlot(fit2000, type=2)
VI_F <- importance(fit2000)[,'MeanDecreaseGini']
par(mar=c(7,4,4,2))
barplot(VI_F/sum(VI_F), las=3)

```

Run one without dryland soy!

```{r rf2000_NoSoy, fig.height=8}
# run a random forest on all 57 variables
set.seed(415)
fit2000.ns <- randomForest(x = pointsNoSoy2[,1:57], # all variable columns 
                    y = as.factor(pointsNoSoy2$classes),
                    ntree = 2000,
                    importance = TRUE)

# look at variable importance
varImpPlot(fit2000.ns)
VI_F.ns <- importance(fit2000.ns)[,'MeanDecreaseGini']
par(mar=c(7,4,4,2))
barplot(VI_F.ns/sum(VI_F.ns), las=3)

```

Plot the error by number of trees
```{r treeError}
plot(fit500)
plot(fit1000)
plot(fit2000)
```



## Annual Random Forest

This dataset doesn't include dryland soy, nor any monthly variables (extracted above in the load data chunk)

```{r rf2000_annual, fig.height=6}
# run a random forest on all 57 variables
set.seed(415)
fit2000.ann <- randomForest(x = annualNoSoy[,1:20], # all variable columns 
                    y = as.factor(annualNoSoy$classes),
                    ntree = 2000,
                    importance = TRUE)

# look at variable importance
varImpPlot(fit2000.ann)
VI_F.ann <- importance(fit2000.ann)[,'MeanDecreaseGini']
par(mar=c(7,4,4,2))
barplot(VI_F.ann/sum(VI_F.ann), las=3)

```


## Play with variable reduction

```{r varReduction}
# run a random forest on all 57 variables
set.seed(412)
test <- varSelRF(xdata = pointsNoSoy2[,1:57], whole.range=F,
                    Class = as.factor(pointsNoSoy2$classes),
                    ntree = 2000)
test

# annual
atest <- varSelRF(xdata = annualNoSoy[,1:20], whole.range=F,
                    Class = as.factor(annualNoSoy$classes),
                    ntree = 2000)
atest
```

