---
title: "Random Forest"
author: "Jill Deines"
date: "January 17, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, cache=TRUE, cache.path='cache/1.18_rrbRandFor/',
               fig.path='figure/1.18_rrbRandFor/')
```

Goal: Run a variable importance test on a random forest of my training points

Update: 2/26/2017 - Classify the MidRep validation points with R's random forest tree and compare with GEE's classification

randomForest package implementation:
 reference implementation based on CART trees
(Breiman, 2001; Liaw and Wiener, 2008)
– for variables of different types: biased in favor of
continuous variables and variables with many categories
(Strobl, Boulesteix, Zeileis, and Hothorn, 2007)


**R Packages Needed**

```{r packages, message=FALSE, echo=TRUE}
library(randomForest)
library(varSelRF)
```

## Load Data
Load training point data used in GEE classification. These were processed in 1.16_trainingPoints_RRB.Rmd, which involved combining 2010 and 2012 points and removing extraneous columns, so that only the variables and a masterType and masterNum columns remained. I wrote out a csv version of the final kml to be used here.

```{r loadData}
# load kml of combined 2010/2012 training points
trainingFolder <- 'C:/Users/deinesji/Dropbox/1PhdJill/hpa/irrigation/data/GIS/training'
trainingName <- '2010_2012_COMBINED_training_12class_v2.csv'

points <- read.csv(paste0(trainingFolder, '/', trainingName))

# re-class the master types into 3 categories: irrigated, dryland crop, noncrop
typeConverter <- data.frame(masterType = unique(points$masterType),
                            classes = c('irrigated','dryland','irrigated','dryland',
                                        'irrigated','dryland','irrigated','dryland',
                                        'irrigated','dryland','noncrop','noncrop'))
points2  <- merge(points, typeConverter)

# remove extra columns
points3 <- points2[ , -which(names(points2) %in% c('masterType','masterNum'))]

# make a dataset without dryland soy
pointsNoSoy <- points2[points2$masterType != 'soy_dryland',]
pointsNoSoy2 <- pointsNoSoy[ , -which(names(pointsNoSoy) %in%
                                        c('masterType','masterNum'))]

# make an annual variable dataset without dryland soy
columnsWanted <- c('EVI_max_14','EVI_range_14','GI_max_14','GI_range_14',
                   'NDVI_max_14','NDVI_range_14','NDWI_max_14','NDWI_range_14',
                   'aridity','b1','b1_1','greenArid','ndwi_gi','pdsi_ann',
                   'pdsi_grow','pr_ann','pr_early','pr_grow','pr_paw','slope_mean',
                   'classes')
annualNoSoy <- pointsNoSoy2[, columnsWanted]
```

## Random Forest

### Monthly random forest
First I do a random forest that mimics my GEE test (500 trees)
```{r rf500, fig.height = 8}
# run a random forest on all 57 variables
set.seed(415)
fit500 <- randomForest(x = points3[,1:57], # all variable columns (excludes classes)
                    y = as.factor(points2$classes),
                    ntree = 500,
                    importance = TRUE)

# look at variable importance
varImpPlot(fit500)
```

That was fast, let's do more trees

```{r rf1000, fig.height = 8}
# run a random forest on all 57 variables
set.seed(415)
fit1000 <- randomForest(x = points3[,1:57], # all variable columns (excludes classes)
                    y = as.factor(points2$classes),
                    ntree = 1000,
                    importance = TRUE)

# look at variable importance
varImpPlot(fit1000)
```

```{r rf2000, fig.height = 8}
# run a random forest on all 57 variables
set.seed(415)
fit2000 <- randomForest(x = points3[,1:57], # all variable columns (excludes classes)
                    y = as.factor(points2$classes),
                    ntree = 2000,
                    importance = TRUE)

# look at variable importance
varImpPlot(fit2000, type=2)
VI_F <- importance(fit2000)[,'MeanDecreaseGini']
par(mar=c(7,4,4,2))
barplot(VI_F/sum(VI_F), las=3)

```

Run one without dryland soy!

```{r rf2000_NoSoy, fig.height=8}
# run a random forest on all 57 variables
set.seed(415)
fit2000.ns <- randomForest(x = pointsNoSoy2[,1:57], # all variable columns 
                    y = as.factor(pointsNoSoy2$classes),
                    ntree = 2000,
                    importance = TRUE)

# look at variable importance
varImpPlot(fit2000.ns)
VI_F.ns <- importance(fit2000.ns)[,'MeanDecreaseGini']
par(mar=c(7,4,4,2))
barplot(VI_F.ns/sum(VI_F.ns), las=3)

```

Plot the error by number of trees
```{r treeError}
plot(fit500)
plot(fit1000)
plot(fit2000)
```



### Annual Random Forest

This dataset doesn't include dryland soy, nor any monthly variables (extracted above in the load data chunk)

```{r rf2000_annual, fig.height=6}
# run a random forest on all 57 variables
set.seed(415)
fit2000.ann <- randomForest(x = annualNoSoy[,1:20], # all variable columns 
                    y = as.factor(annualNoSoy$classes),
                    ntree = 2000,
                    importance = TRUE)

# look at variable importance
varImpPlot(fit2000.ann)
VI_F.ann <- importance(fit2000.ann)[,'MeanDecreaseGini']
par(mar=c(7,4,4,2))
barplot(VI_F.ann/sum(VI_F.ann), las=3)

```


### Play with variable reduction

```{r varReduction}
# run a random forest on all 57 variables
set.seed(412)
test <- varSelRF(xdata = pointsNoSoy2[,1:57], whole.range=F,
                    Class = as.factor(pointsNoSoy2$classes),
                    ntree = 2000)
test

# annual
atest <- varSelRF(xdata = annualNoSoy[,1:20], whole.range=F,
                    Class = as.factor(annualNoSoy$classes),
                    ntree = 2000)
atest
```

## Make Predictions
Use the R-generated random forest to make predictions on the Middle Republican validation points. Mimic the GEE rf classification, so 500 trees? 

This is for the monthly classifier.

### predict
Use the validation points that have sampled the 57 input variables from the 01.1_ValidationPointsSample script in my GEE 'validation' folder

```{r predict}
# load points and variable data
vpointFolder <- 'C:/Users/deinesji/Google Drive/GEE_validation/validationSets'
p2007 <- read.csv(paste0(vpointFolder,'/2007_test3_variables_midRep_all.csv'))
p2010 <- read.csv(paste0(vpointFolder,'/2010_test3_variables_midRep_all.csv'))

# par down dfs
unique(p2007$certainty)
p2007 <- p2007[ , -which(names(p2007) %in% c('.geo'))]
p2010 <- p2010[ , -which(names(p2010) %in% c('.geo'))]


# run the random forest (monthly, no soy, all 57 variables)
# run a random forest on all 57 variables
set.seed(415)
fit500.ns <- randomForest(x = pointsNoSoy2[,1:57], # all variable columns 
                    y = as.factor(pointsNoSoy2$classes),
                    ntree = 500,
                    importance = TRUE)

# look at variable importance
varImpPlot(fit500.ns)

# run predictions/classification
p2007$randFor <- predict(fit500.ns, newdata = p2007, type='response')
p2010$randFor <- predict(fit500.ns, newdata = p2010, type='response')
```

### assess
Use the accuracy assessment exports from GEE validation to compare results between GEE and Rrrr

```{r compare}
# load GEE accuracy tables
accuDir <- 'C:/Users/deinesji/Google Drive/GEE_tableExports/Accuracy_PointsforConfusionTables_test1_clean'
g2007 <- read.csv(paste0(accuDir,
        '/2007_MidRep60_confusionData_RRB_test3_randFor_interannual_maskFixed_allClasses.csv'))
g2010 <- read.csv(paste0(accuDir,
        '/2010_MidRep60_confusionData_RRB_test3_randFor_interannual_maskFixed_allClasses.csv'))

g2007 <- g2007[ , -which(names(g2007) %in% c('.geo'))]
g2010 <- g2010[ , -which(names(g2010) %in% c('.geo'))]

# extract important columns from r dataset
rColNamesWanted <- c('system.index','class','classNum','randFor')
p2007.less <- p2007[,rColNamesWanted]
p2010.less <- p2010[,rColNamesWanted]

# combine
geeR.2007 <- merge(g2007, p2007.less, by = 'system.index',all=F)
geeR.2010 <- merge(g2010, p2010.less, by = 'system.index',all=F)
# make sure the system index is an ok column to merge on
sum(!(geeR.2007$classNum.x == geeR.2007$classNum.y))

# convert GEE's classifications from numbers to words
key7 <- data.frame(b2007 = c(0,1,2),
                  gee_rf = c('dryland','irrigated','noncrop'))

key10 <- data.frame(b2010 = c(0,1,2),
                  gee_rf = c('dryland','irrigated','noncrop'))

geeR.2007 <- merge(geeR.2007, key7)
geeR.2010 <- merge(geeR.2010, key10)

# quick check for matches between gee and R
sum(!(geeR.2007$gee_rf == geeR.2007$randFor))
sum(!(geeR.2010$gee_rf == geeR.2010$randFor))

# give percentages
(nrow(geeR.2007) - sum(!(geeR.2007$gee_rf == geeR.2007$randFor))) / nrow(geeR.2007)
(nrow(geeR.2010) - sum(!(geeR.2010$gee_rf == geeR.2010$randFor))) / nrow(geeR.2010)
```

Not awful, but check to see if most of the confusion is between noncrop/dryland

```{r compare2class}
# make keys to convert classifications from 2 classes to 1
key2class.1 = data.frame(gee_rf = c('dryland','irrigated','noncrop'),
                         gee_rf2 = c('nonirr','irrigated','nonirr'))
key2class.2 = data.frame(randFor = c('dryland','irrigated','noncrop'),
                         randFor2 = c('nonirr','irrigated','nonirr'))

geeR.2007 <- merge(geeR.2007, key2class.1)
geeR.2007 <- merge(geeR.2007, key2class.2)

geeR.2010 <- merge(geeR.2010, key2class.1)
geeR.2010 <- merge(geeR.2010, key2class.2)

# quick check for matches between gee and R
sum(!(geeR.2007$gee_rf2 == geeR.2007$randFor2))
sum(!(geeR.2010$gee_rf2 == geeR.2010$randFor2))

# give percentages
(nrow(geeR.2007) - sum(!(geeR.2007$gee_rf2 == geeR.2007$randFor2))) / nrow(geeR.2007)
(nrow(geeR.2010) - sum(!(geeR.2010$gee_rf2 == geeR.2010$randFor2))) / nrow(geeR.2010)
```

That does improve the "consistency" between programs. Or likely, random seeds.

Check to see which is more accurate?

```{r compareAccuracy}

```

