---
title: "Validation Points - Others"
author: "Jill Deines"
date: "Friday, November 04, 2016"
output: 
  html_document:
    toc: yes
---

Goal: process validation data sets for Kansas from Brian Wardlow (UNL).

Note: This script is ridiculously poorly organized, don't judge.

```{r knitrOpts, echo=FALSE}
library(knitr)
opts_chunk$set(cache=TRUE, cache.path='cache/1.3_vOthers/',
               fig.path='figure/1.3_vOthers/')
```

**R Packages Needed**

```{r packages, message=FALSE, echo=TRUE}
library(rgdal)
library(raster)
library(rgeos)

library(tidyr)
library(ggplot2)

library(cluster) 
library(fpc)

# source project helper functions
source('0.50_helperFunctions.R')
```

## Points to KML for Inspection
Points were provided in .csv format. In an email, Brian Wardlow said that the coordinates are in Lambert Azimuthal Equal Area and Sphere of the Radius (spheroid and datum). I translated this as http://spatialreference.org/ref/sr-org/80/ and Wardlow said those specs were correct.

Dataset includes 2,179 field sites (some outside of the HPA) from 2001, and include crop type and irrigation status.


```{r rawToKml}
dataDir <- 'C:/Users/deinesji/Dropbox/1PhdJill/hpa/irrigation/data/GIS/validationFromOthers'
#dataDir <- 'S:/Users/deinesji/HPA/gis/rsIrrigation/validationOthers'
dataFile <- 'Kansas_field_sites.csv'

# load a contextual shapefile to confirm projection
gisDir <- 'C:/Users/deinesji/Dropbox/1PhdJill/hpa/irrigation/data/GIS/kml'
#gisDir <- 'S:/Users/deinesji/HPA/gis'
rrb <- readOGR(gisDir, 'BigExtent_RRB_RRCA_buff', verbose=F))

# load data
points <- read.csv(paste(dataDir,dataFile, sep='/'))

# possible proj4 string for Sphere_ARC_INFO_Lambert_Azimuthal_Equal_Area
# from http://spatialreference.org/ref/sr-org/80/
proj <- '+proj=laea +lat_0=45 +lon_0=-100 +x_0=0 +y_0=0 +a=6370997 +b=6370997 +units=m +no_defs' 

coordinates(points) <- ~X + Y
proj4string(points) <- proj

# export for projection test in ArcGIS
#writeOGR(points, sDir, 'Kansas_field_sites', driver='ESRI Shapefile')

points2 <- spTransform(points, CRS(proj4string(rrb)))

# plot
plot(rrb)
plot(points2, add=T)

# export points for further investigation
points.ll <- spTransform(points, CRS("+proj=longlat +datum=WGS84"))
#writeOGR(points.ll, paste0(sDir, '/', 'Kansas_field_sites.kml'),
            layer = 'points', driver = 'KML')
```

## Clean Bad Point Locations, RRB
In GEE and ArcGIS, ~15% of points are clearly located adjacent to the intended field. I made a new point dataset at points with "wrong place" or suspected "wrong class". I then sampled annual greenness indices at all points for outlier analysis.

Here, I see if outlier detection can deal with these points in an unbiased and justified manner.

### load and clip points

```{r cleanPoints_RRB_load}
# load RRB boundary
gisDir <- 'C:/Users/deinesji/Dropbox/1PhdJill/hpa/irrigation/data/GIS/kml'
#gisDir <- 'S:/Users/deinesji/HPA/gis'
rrb <- readOGR(paste0(gisDir,'/BigExtent_RRB_RRCA_buff.kml'), 'boundExtent', verbose=F)

# load datasets exported from GEE
dataDir <- 'C:/Users/deinesji/Dropbox/1PhdJill/hpa/irrigation/data/GIS/validationFromOthers'
dataFile1 <- 'Kansas_field_sites_sampledGreenness.csv'
dataFile2 <- 'Kansas_field_sites_flagged.csv'

# load csv's and convert to spdf
allpoints <- csvToSpdf(paste(dataDir,dataFile1, sep='/')) # sourced function
flagged <- csvToSpdf(paste(dataDir,dataFile2, sep='/'))

# proj strings identical
rrb <- spTransform(rrb, CRS(proj4string(allpoints)))

# clip points to those inside RRB
rrbpoints0 <- allpoints[rrb,]
flagged <- flagged[rrb,]

# get the actual percentage of suspected bad points
sum(flagged$type == 'wrongPlace')/nrow(rrbpoints0)

plot(rrb)
plot(rrbpoints0, add=T)
plot(flagged, col='red', add=T)
```

### Flag suspect points
Points I suspect are in the wrong place or the wrong type.

```{r flagPoints}
# project to meters
rrbMeters <- spTransform(rrbpoints0, CRS('+proj=aea +lat_1=29.5 +lat_2=45.5 +lat_0=37.5 +lon_0=-96 +x_0=0 +y_0=0 +ellps=GRS80 +datum=NAD83 +units=m +no_defs '))
flagMeters <- spTransform(flagged, CRS('+proj=aea +lat_1=29.5 +lat_2=45.5 +lat_0=37.5 +lon_0=-96 +x_0=0 +y_0=0 +ellps=GRS80 +datum=NAD83 +units=m +no_defs '))

# find distances between all points
distPairs <- as.data.frame(gDistance(flagMeters, rrbMeters, byid = T))
distPairs$cellnum <- 1:nrow(distPairs)

# get paired points
pointBuddies <- data.frame(flagged = 1:nrow(flagMeters), 
                           rrb = NA, 
                           type = flagMeters$type)
for (i in 1:nrow(flagMeters)) {
  pointBuddies[i,'rrb'] <- distPairs[which.min(distPairs[,i]),'cellnum']
}
  
# add flagged status to rrb points
rrbpoints0$ind <- 1:nrow(rrbpoints0)
rrbpoints <- merge(rrbpoints0, pointBuddies[,c('rrb','type')], 
                   by.x = 'ind', by.y = 'rrb', all.X = T)
  
manualClean <- rrbpoints[is.na(rrbpoints$type),]
```

### format data

```{r cleanPoints_rrb_format}
# add a point id column
rrbpoints$ID <- 1:nrow(rrbpoints)
# drop nonsense columns
rrbpoints <- rrbpoints[,-which(names(rrbpoints) %in% c('description','name'))]
# split crop/water status to two columns
rrbpoints$irrigated <- sapply(strsplit(rrbpoints$Crop_type,"\\-"), `[`, 2)
rrbpoints$crop <- sapply(strsplit(rrbpoints$Crop_type,"\\-"), `[`, 1)

# convert spdf to long df
rrblong <- gather(as.data.frame(rrbpoints), key = Index, value = value, EVI_max_14:NDWI_max_14)

# manual rrb clean ------------------
manualClean$ID <- 1:nrow(manualClean)
# drop nonsense columns
manualClean <- manualClean[,-which(names(manualClean) %in% c('description','name'))]
# split crop/water status to two columns
manualClean$irrigated <- sapply(strsplit(manualClean$Crop_type,"\\-"), `[`, 2)
manualClean$crop <- sapply(strsplit(manualClean$Crop_type,"\\-"), `[`, 1)

# convert spdf to long df
manualLong <- gather(as.data.frame(manualClean), key = Index, value = value, 
                     EVI_max_14:NDWI_max_14)



# all KS points ---------------------
allpoints$ID <- 1:nrow(allpoints)
# drop nonsense columns
allpoints <- allpoints[,-which(names(allpoints) %in% c('description','name'))]
# split crop/water status to two columns
allpoints$irrigated <- sapply(strsplit(allpoints$Crop_type,"\\-"), `[`, 2)
allpoints$crop <- sapply(strsplit(allpoints$Crop_type,"\\-"), `[`, 1)

# convert spdf to long df
alllong <- gather(as.data.frame(allpoints), key = Index, value = value, EVI_max_14:NDWI_max_14)
```


### exploratory plots: histograms by crop

```{r rrb_histsCrop}
# GI, RRB
ggplot(data = rrblong[rrblong$Index == 'GI_max_14',], 
       aes(x = value, fill=irrigated)) +
       geom_histogram(alpha=.5, position='identity') +
       facet_wrap(~crop, nrow = 3) + 
       ylab('count') + xlab('Greenness Index') + theme_bw() +
       ggtitle("All RRB Points, GI")

# GI, RRB manual clean
ggplot(data = manualLong[manualLong$Index == 'GI_max_14',], 
       aes(x = value, fill=irrigated)) +
       geom_histogram(alpha=.5, position='identity') +
       facet_wrap(~crop, nrow = 3) + 
       ylab('count') + xlab('Greenness Index') + theme_bw() +
       ggtitle("manual clean RRB Points, GI")

# NDWI, RRB
ggplot(data = rrblong[rrblong$Index == 'NDWI_max_14',], 
       aes(x = value, fill=irrigated)) +
       geom_histogram(alpha=.5, position='identity') +
       facet_wrap(~crop, nrow = 3) + 
       ylab('count') + xlab('NDWI') + theme_bw() +
       ggtitle("All RRB Points, NDWI")

# NDVI, RRB
ggplot(data = rrblong[rrblong$Index == 'NDVI_max_14',], 
       aes(x = value, fill=irrigated)) +
       geom_histogram(alpha=.5, position='identity') +
       facet_wrap(~crop, nrow = 3) + 
       ylab('count') + xlab('NDVI') + theme_bw() +
       ggtitle("All RRB Points, NDVI")

# EVI, RRB
ggplot(data = rrblong[rrblong$Index == 'EVI_max_14',], 
       aes(x = value, fill=irrigated)) + xlim(0,1) +
       geom_histogram(alpha=.5, position='identity') +
       facet_wrap(~crop, nrow = 3) + 
       ylab('count') + xlab('EVI') + theme_bw() +
       ggtitle("All RRB Points, EVI")
```


```{r ks_histsCrop}
# GI, KS
ggplot(data = alllong[alllong$Index == 'GI_max_14',], 
       aes(x = value, fill=irrigated)) +
       geom_histogram(alpha=.5, position='identity') +
       facet_wrap(~crop, nrow = 3) + 
       ylab('count') + xlab('Greenness Index') + theme_bw() +
       ggtitle("All KS Points, GI")

# NDWI, KS
ggplot(data = alllong[alllong$Index == 'NDWI_max_14',], 
       aes(x = value, fill=irrigated)) +
       geom_histogram(alpha=.5, position='identity') +
       facet_wrap(~crop, nrow = 3) + 
       ylab('count') + xlab('NDWI') + theme_bw() +
       ggtitle("All KS Points, NDWI")

# NDVI, KS
ggplot(data = alllong[alllong$Index == 'NDVI_max_14',], 
       aes(x = value, fill=irrigated)) +
       geom_histogram(alpha=.5, position='identity') +
       facet_wrap(~crop, nrow = 3) + 
       ylab('count') + xlab('NDVI') + theme_bw() +
       ggtitle("All KS Points, NDVI")

# EVI, KS
ggplot(data = alllong[alllong$Index == 'EVI_max_14',], 
       aes(x = value, fill=irrigated)) + xlim(0,1) +
       geom_histogram(alpha=.5, position='identity') +
       facet_wrap(~crop, nrow = 3) + 
       ylab('count') + xlab('EVI') + theme_bw() +
       ggtitle("All KS Points, EVI")
```

### exploratory plots: 

```{r plots}
dataset = as.data.frame(rrbpoints)

ggplot(data=dataset,
       aes(x = GI_max_14, y = NDWI_max_14, color=irrigated)) +
       geom_point() +
       facet_wrap(~crop, nrow = 3) + 
       ylab('NDWI') + xlab('GI') + theme_bw() +
       ggtitle("GI vs NDWI")

# climate variables
ggplot(data=dataset,
       aes(x = GI_max_14, y = pdsi_grow, color=irrigated)) +
       geom_point() +
       facet_wrap(~crop, nrow = 3) + 
       ylab('PDSI') + xlab('GI') + theme_bw() +
       ggtitle("GI vs PDSI, all RRB")

ggplot(data=dataset,
       aes(x = NDWI_max_14, y = pdsi_grow, color=irrigated)) +
       geom_point() +
       facet_wrap(~crop, nrow = 3) + 
       ylab('PDSI') + xlab('NDWI') + theme_bw() +
       ggtitle("NDWI vs PDSI, all RRB")

ggplot(data=dataset,
       aes(x = GI_max_14, y = pr_grow, color=irrigated)) +
       geom_point() +
       facet_wrap(~crop, nrow = 3) + 
       ylab('growing precip') + xlab('GI') + theme_bw() +
       ggtitle("GI vs precip, all RRB")

ggplot(data=dataset,
       aes(x = NDWI_max_14, y = pr_grow, color=irrigated)) +
       geom_point() +
       facet_wrap(~crop, nrow = 3) + 
       ylab('growing precip') + xlab('NDWI') + theme_bw() +
       ggtitle("NDWI vs precip, all RRB")

```

## Outliers

### Quantiles

```{r quantilesFun}
# function from an R Blogger Klodian Dhana
outlierKD <- function(dt, var) {
     var_name <- eval(substitute(var),eval(dt))
     na1 <- sum(is.na(var_name))
     m1 <- mean(var_name, na.rm = T)
     par(mfrow=c(2, 2), oma=c(0,0,3,0))
     boxplot(var_name, main="With outliers")
     hist(var_name, main="With outliers", xlab=NA, ylab=NA)
     outlier <- boxplot.stats(var_name)$out
     mo <- mean(outlier)
     var_name <- ifelse(var_name %in% outlier, NA, var_name)
     boxplot(var_name, main="Without outliers")
     hist(var_name, main="Without outliers", xlab=NA, ylab=NA)
     title("Outlier Check", outer=TRUE)
     na2 <- sum(is.na(var_name))
     cat("Outliers identified:", na2 - na1, "n")
     cat("Propotion (%) of outliers:", round((na2 - na1) / sum(!is.na(var_name))*100, 1), "n")
     cat("Mean of the outliers:", round(mo, 2), "n")
     m2 <- mean(var_name, na.rm = T)
     cat("Mean without removing outliers:", round(m1, 2), "n")
     cat("Mean if we remove outliers:", round(m2, 2), "n")
     
     # remove outliers
      dt[as.character(substitute(var))] <- invisible(var_name)
      assign(as.character(as.list(match.call())$dt), dt, envir = .GlobalEnv)
      cat("Outliers successfully removed", "n")
      return(invisible(dt))
}

```


Use all 4 indices to reduce by IQR outliers

```{r quantiles}
## Irrigated Corn --------------------------------------------------------------
# fun outlier function by 
rrbCornI <- rrbpoints@data[rrbpoints@data$Crop_type == 'Corn-I',]
# remove outliers for corn
outlierKD(rrbCornI, EVI_max_14)
outlierKD(rrbCornI, NDVI_max_14)
outlierKD(rrbCornI, GI_max_14)
outlierKD(rrbCornI, NDWI_max_14)

## Dryland Corn --------------------------------------------------------------
rrbCornN <- rrbpoints@data[rrbpoints@data$Crop_type == 'Corn-NI',]
# remove outliers for corn
outlierKD(rrbCornN, EVI_max_14)
outlierKD(rrbCornN, NDVI_max_14)
outlierKD(rrbCornN, GI_max_14)
outlierKD(rrbCornN, NDWI_max_14)

# Irrigated Alfalfa ---------------------------------------------
rrbAlfI <- rrbpoints@data[rrbpoints@data$Crop_type == 'Alfalfa-I',]
outlierKD(rrbAlfI, EVI_max_14)
outlierKD(rrbAlfI, NDVI_max_14)
outlierKD(rrbAlfI, GI_max_14)
outlierKD(rrbAlfI, NDWI_max_14)

# Dryland Alfalfa -----------------------------------------------------
rrbAlfN <- rrbpoints@data[rrbpoints@data$Crop_type == 'Alfalfa-NI',]
outlierKD(rrbAlfN, EVI_max_14)
outlierKD(rrbAlfN, NDVI_max_14)
outlierKD(rrbAlfN, GI_max_14)
outlierKD(rrbAlfN, NDWI_max_14)

# Irrigated Soy -----------------------------------------------------
rrbSoyI <- rrbpoints@data[rrbpoints@data$Crop_type == 'Soybeans-I',]
outlierKD(rrbSoyI, EVI_max_14)
outlierKD(rrbSoyI, NDVI_max_14)
outlierKD(rrbSoyI, GI_max_14)
outlierKD(rrbSoyI, NDWI_max_14)

# Dryland Soy -----------------------------------------------------
rrbSoyN <- rrbpoints@data[rrbpoints@data$Crop_type == 'Soybeans-NI',]
outlierKD(rrbSoyN, EVI_max_14)
outlierKD(rrbSoyN, NDVI_max_14)
outlierKD(rrbSoyN, GI_max_14)
outlierKD(rrbSoyN, NDWI_max_14)

# Irrigated wheat -----------------------------------------------------
rrbWheatI <- rrbpoints@data[rrbpoints@data$Crop_type == 'Wheat-I',]
outlierKD(rrbWheatI, EVI_max_14)
outlierKD(rrbWheatI, NDVI_max_14)
outlierKD(rrbWheatI, GI_max_14)
outlierKD(rrbWheatI, NDWI_max_14)

# Dryland wheat -----------------------------------------------------
rrbWheatN <- rrbpoints@data[rrbpoints@data$Crop_type == 'Wheat-NI',]
outlierKD(rrbWheatN, EVI_max_14)
outlierKD(rrbWheatN, NDVI_max_14)
outlierKD(rrbWheatN, GI_max_14)
outlierKD(rrbWheatN, NDWI_max_14)

# Irrigated sorghum -----------------------------------------------------
rrbSorghumI <- rrbpoints@data[rrbpoints@data$Crop_type == 'Sorghum-I',]
outlierKD(rrbSorghumI, EVI_max_14)
outlierKD(rrbSorghumI, NDVI_max_14)
outlierKD(rrbSorghumI, GI_max_14)
outlierKD(rrbSorghumI, NDWI_max_14)

# Dryland sorghum -----------------------------------------------------
rrbSorghumN <- rrbpoints@data[rrbpoints@data$Crop_type == 'Sorghum-NI',]
outlierKD(rrbSorghumN, EVI_max_14)
outlierKD(rrbSorghumN, NDVI_max_14)
outlierKD(rrbSorghumN, GI_max_14)
outlierKD(rrbSorghumN, NDWI_max_14)

# fallow ---------------------------------------------------------
rrbFallow <- rrbpoints@data[rrbpoints@data$Crop_type == 'Fallow',]
outlierKD(rrbFallow, EVI_max_14)
outlierKD(rrbFallow, NDVI_max_14)
outlierKD(rrbFallow, GI_max_14)
outlierKD(rrbFallow, NDWI_max_14)


# reassemble and process ----------------------------------------
rrbpoints2 <- rbind(rrbCornI, rrbCornN, rrbAlfI, rrbAlfN, rrbWheatI,
                    rrbWheatN, rrbSoyI, rrbSoyN, rrbSorghumI, rrbSorghumN,
                    rrbFallow)
# make a column with NA anywhere an outlier removed
for (i in 1:nrow(rrbpoints2)){
  rrbpoints2$outlier[i] <- sum(rrbpoints2[i,c('EVI_max_14','GI_max_14','NDVI_max_14','NDWI_max_14')])
}
  
# remove na's
rrbpoints3 <- rrbpoints2[!is.na(rrbpoints2$outlier),]

# convert spdf to long df
rrbOuts <- gather(rrbpoints3, key = Index, value = value, EVI_max_14:NDWI_max_14)

ggplot(data = rrbOuts[rrbOuts$Index == 'GI_max_14',], 
       aes(x = value, fill=irrigated)) +
       geom_histogram(alpha=.5, position='identity') +
       facet_wrap(~crop, nrow = 3) + 
       ylab('count') + xlab('Greenness Index') + theme_bw() +
       ggtitle("RRB Points, no Outliers, GI")

```

Use indices that best match flagged points for each crop category

```{r quantilesMinimalist}
## Irrigated Corn --------------------------------------------------------------
# fun outlier function by 
rrbCornI <- rrbpoints@data[rrbpoints@data$Crop_type == 'Corn-I',]
# remove outliers for corn
outlierKD(rrbCornI, EVI_max_14)
outlierKD(rrbCornI, NDWI_max_14)

## Dryland Corn --------------------------------------------------------------
rrbCornN <- rrbpoints@data[rrbpoints@data$Crop_type == 'Corn-NI',]
# no points flagged

# Irrigated Alfalfa ---------------------------------------------
rrbAlfI <- rrbpoints@data[rrbpoints@data$Crop_type == 'Alfalfa-I',]
outlierKD(rrbAlfI, NDWI_max_14)

# Dryland Alfalfa -----------------------------------------------------
rrbAlfN <- rrbpoints@data[rrbpoints@data$Crop_type == 'Alfalfa-NI',]
outlierKD(rrbAlfN, GI_max_14)

# Irrigated Soy -----------------------------------------------------
rrbSoyI <- rrbpoints@data[rrbpoints@data$Crop_type == 'Soybeans-I',]
outlierKD(rrbSoyI, NDWI_max_14)

# Dryland Soy -----------------------------------------------------
rrbSoyN <- rrbpoints@data[rrbpoints@data$Crop_type == 'Soybeans-NI',]
outlierKD(rrbSoyN, EVI_max_14)

# Irrigated wheat -----------------------------------------------------
rrbWheatI <- rrbpoints@data[rrbpoints@data$Crop_type == 'Wheat-I',]
outlierKD(rrbWheatI, NDWI_max_14)

# Dryland wheat -----------------------------------------------------
rrbWheatN <- rrbpoints@data[rrbpoints@data$Crop_type == 'Wheat-NI',]
# one flagged, none picked up by indices

# Irrigated sorghum -----------------------------------------------------
rrbSorghumI <- rrbpoints@data[rrbpoints@data$Crop_type == 'Sorghum-I',]
# none flagged

# Dryland sorghum -----------------------------------------------------
rrbSorghumN <- rrbpoints@data[rrbpoints@data$Crop_type == 'Sorghum-NI',]
# none flagged

# fallow ---------------------------------------------------------
rrbFallow <- rrbpoints@data[rrbpoints@data$Crop_type == 'Fallow',]
outlierKD(rrbFallow, GI_max_14)

# reassemble and process ----------------------------------------
rrbpoints2 <- rbind(rrbCornI, rrbCornN, rrbAlfI, rrbAlfN, rrbWheatI,
                    rrbWheatN, rrbSoyI, rrbSoyN, rrbSorghumI, rrbSorghumN,
                    rrbFallow)
# make a column with NA anywhere an outlier removed
for (i in 1:nrow(rrbpoints2)){
  rrbpoints2$outlier[i] <- sum(rrbpoints2[i,c('EVI_max_14','GI_max_14','NDVI_max_14','NDWI_max_14')])
}
  
# remove na's
rrbpoints3 <- rrbpoints2[!is.na(rrbpoints2$outlier),]

# convert spdf to long df
rrbOuts <- gather(rrbpoints3, key = Index, value = value, EVI_max_14:NDWI_max_14)

ggplot(data = rrbOuts[rrbOuts$Index == 'GI_max_14',], 
       aes(x = value, fill=irrigated)) +
       geom_histogram(alpha=.5, position='identity') +
       facet_wrap(~crop, nrow = 3) + 
       ylab('count') + xlab('Greenness Index') + theme_bw() +
       ggtitle("RRB Points, minimal Outliers, GI")

```

### K Means Aborted attempt
Functions to do outlier tests. Includes k-means clustering..

```{r kmeans_fun}
# function to visualize cluster numbers
clusterPlot <- function(mydata){
  # Determine number of clusters
  wss <- (nrow(mydata)-1)*sum(apply(mydata,2,var))
  for (i in 2:15) wss[i] <- sum(kmeans(mydata, 
      centers=i)$withinss)
  plot(1:15, wss, type="b", xlab="Number of Clusters",
    ylab="Within groups sum of squares")
}

# run kmeans
kmeansTest <- function(mydata, numClusters){
    # K-Means Cluster Analysis
  fit <- kmeans(mydata, numClusters) # 5 cluster solution
  # get cluster means 
  aggregate(mydata,by=list(fit$cluster),FUN=mean)
  # append cluster assignment
  dataout <- data.frame(mydata, fit$cluster)
  
  # Cluster Plot against 1st 2 principal components
  # vary parameters for most readable graph
  clusplot(dataout, fit$cluster, color=TRUE, shade=TRUE, 
      labels=2, lines=0)
  
  return(dataout)
}

```


```{r kmeans}
# irrigated corn ------------------------------------------
# extract corn points and standardize variables
rrbCornI <- rrbpoints@data[rrbpoints@data$Crop_type == 'Corn-I',c(3:8)]
rrbCornI <- scale(rrbCornI)
# run cluster analysis
clusterPlot(rrbCornI)
set.seed(10)
cornI <- kmeansTest(rrbCornI, 5)
# put fits back with original data
cornI2 <- data.frame(rrbpoints@data[rrbpoints@data$Crop_type == 'Corn-I',], 
                     fit.cluster = cornI$fit.cluster)
# sort by fit.cluster
cornI2 <- cornI2[order(cornI2$fit.cluster),]


# irrigated alfalfa ------------------------------------------
# extract corn points and standardize variables
rrbAlfI <- rrbpoints@data[rrbpoints@data$Crop_type == 'Alfalfa-I',c(3:8)]
rrbAlfI <- scale(rrbAlfI)
# run cluster analysis
clusterPlot(rrbAlfI)
set.seed(10)
kout <- kmeansTest(rrbAlfI, 3)
# put fits back with original data
rrbAlfI2 <- data.frame(rrbpoints@data[rrbpoints@data$Crop_type == 'Alfalfa-I',], 
                     fit.cluster = kout$fit.cluster)
# sort by fit.cluster
rrbAlfI2 <- rrbAlfI2[order(rrbAlfI2$fit.cluster),]

# soy irrigated ------------------------------------------
# extract and standardize variables
rrbSoyI <- rrbpoints@data[rrbpoints@data$Crop_type == 'Soybeans-I',c(3:8)]
rrbSoyI <- scale(rrbSoyI)
# run cluster analysis
clusterPlot(rrbSoyI)
set.seed(10)
kout <- kmeansTest(rrbSoyI, 7)
# put fits back with original data
rrbSoyI2 <- data.frame(rrbpoints@data[rrbpoints@data$Crop_type == 'Soybeans-I',], 
                     fit.cluster = kout$fit.cluster)
# sort by fit.cluster
rrbSoyI2 <- rrbSoyI2[order(rrbSoyI2$fit.cluster),]



```




