---
title: "NASS vs Remote Sensing Validation"
author: "Jill Deines"
date: "April 5, 2017"
output: 
  html_document:
    toc: yes
---

Goal: Compare irrigated acreages between remote sensing products and USDA NASS Ag Census Data. NASS data acquired through the Quick Stats API.

Re-run on 1/25/17 with updated NASS cleaning function (removing double counting in wheat, cotton, and sunflower). Grass/legumes and hay still a bit off. 

update 2/21/2017: adding USGS Waterdata county acreage for years 2000, 2005, 2010; used only counties 100% within boundary

update 2/24/2017 - uses the regional stats files instead of the county files

update/new file 4/5/2017: run with cleaned maps produced by the annual classifier


```{r knitrOpts, echo=FALSE}
library(knitr)
opts_chunk$set(cache=FALSE, cache.path='cache/2.13_NASS_rf/',
               fig.path='figure/2.14_NASS_rf/')
```

**R Packages Needed**

* `httr`: Vignette at https://cran.r-project.org/web/packages/httr/vignettes/quickstart.html

```{r packages, message=FALSE, echo=TRUE}
# for API interaction
library(httr) 
library(jsonlite)
library(maps)  # to translate state fp codes to 2 letter abbrevs
library(rgdal)
library(tigris)
library(RColorBrewer)

# for data manipulation
library(tidyr)
library(ggplot2)

# my functions (to be packaged)
source('functions/getNASS.R')
source('functions/cleanNassCensusCounty.R')
source('functions/stat_smooth_func.R')
```


## Load county data
In GEE, I calculated the number of cells of each classification type in each county overlapping the RRB/RRCA boundary.

Here, I load this dataset from yearly files, add a year attribute, and converting cell counts to areas (both retained).


```{r loadGee}
# years with county level datasets
yearsDesired <- c(2000, 2002, 2005, 2007, 2010, 2012)
cellArea <- 30*30 # m^2

countyDir <- 'C:/Users/deinesji/Google Drive/GEE_tableExports/RRB_test4_regional_county_Stats'

# get filenames
countyfiles <- list.files(countyDir, pattern="*test4_randFor_interannual_plusAncillary.csv")

# subset to years wanted
yearsAvailable <- as.numeric(substr(countyfiles, start =1, stop = 4))
wanted <- yearsAvailable %in% yearsDesired
countyfiles <- countyfiles[wanted]

# read in to a single dataframe
countyAll = do.call(rbind, lapply(countyfiles, function(x) {
             # print(x)
              csv0 <- read.csv(paste0(countyDir,'/',x), stringsAsFactors = FALSE)
              # extract county regions only
              csv <- csv0[grepl('county',csv0$masterID),]
              # make columns
              csv$year <- as.numeric(substr(x, start=1,stop=4))
              csv$fips5 <- substr(csv$masterID, start=11,stop=15)
              csv$fips3 <- substr(csv$masterID, start=13,stop=15)
              csv$state2 <- substr(csv$masterID, start=11,stop=12)
              csv <- csv[,c('fips5','state2','fips3','X0','X1','masterID','year')]
            }))

# subset counties -------------------
# get kml with county proportion in bounds
countyShape <- readOGR('C:/Users/deinesji/Dropbox/1PhdJill/hpa/irrigation/data/GIS/kml/shapefile',
            'Counties_RRCAextent_Clip', verbose=F)
# retrieve list of fips within bounds
fips.in <- as.character(countyShape[countyShape$proportion == 1,]$fips5)

# keep counties 100% within the rrca
countyAll <- countyAll[countyAll$fips5 %in% fips.in,]

# convert cell count to area
countyAll$irrigated_m2 <- countyAll$X1 * cellArea
countyAll$irrArea.km2 <- countyAll$irrigated_m2 / 1000000

# # convert dryland cell count to area
# countyAll$dryland_m2 <- countyAll$X0 * cellArea
# countyAll$dryland_km2 <- countyAll$dryland_m2 / 1000000
# 
# # convert cropland area to km^2 for RS
# # convert cell count to area
# countyAll$cropland_m2 <- (countyAll$X1+countyAll$X0) * cellArea
# countyAll$cropland_km2 <- countyAll$cropland_m2 / 1000000

# add a state 2-letter abbrev column
# fill in state abbrevs column based on FIPS
data(state.fips)  # from maps package
state.fips$STATEFP <- sprintf("%02d",state.fips$fips) # add leading zero
countyAll <- merge(countyAll, state.fips[,c('abb','STATEFP')], 
                   by.x = 'state2', by.y = 'STATEFP',all.y=F)

# # write out csv for Anthony to use
# anthony <- countyAll[countyAll$year == 2002,c('state2','fips5','fips3','abb')]
# outdir <- 'C:/Users/deinesji/Dropbox/1PhdJill/hpa/irrigation/data/nass'
# outfile3 <- 'NASS_countiesToGet_forADK.csv'
# write.csv(anthony, paste0(outdir, '/', outfile3), row.names=F)
```

## Get and Clean NASS Data
From the NASS Quick Stats API, using county list generated above

re-run on 1/24/2017 to get all counties 90% or greater within RRB/RRCA bound.
On this date, also retrieved data for all counties in the HPA using 2.00_NASS_countyStats.Rmd, in order to safeguard data.

### Get Data 
And write out to not rely on NASS API

* API website: http://quickstats.nass.usda.gov/api
* My API key: 28EAA9E6-8060-34A4-981A-B2ED4692228A

```{r getNASS, eval=FALSE}
# set up variables for data calls
apiKey <- '28EAA9E6-8060-34A4-981A-B2ED4692228A'
program = 'CENSUS'
sector <- 'CROPS'
group <- 'FIELD CROPS'
aggregation <- 'COUNTY'

# get just 1 set of counties (so for 1 year)
counties <- countyAll[countyAll$year == unique(countyAll$year)[3],]

# make empty list to populate, and loop through counties
nass.list <- list()
for(i in 1:nrow(counties)){
  nass.list[[i]] <- getNASS(apiKey, program, sector, group, aggregation, 
                              state = counties$abb[i],
                              county = counties$fips3[i])
}

# convert list of data frames to 1 giant dataframe
nass.df <- do.call("rbind",nass.list)

# write out raw data
outdir <- 'C:/Users/deinesji/Dropbox/1PhdJill/hpa/irrigation/data/nass'
outfile <- 'NASS_FieldCrops_County_RRB_raw_90_47counties.csv'
write.csv(nass.df, paste0(outdir, '/', outfile), row.names=F)
```


### clean data
load raw nass download file, clean it up. also has data for 1997 but not used here

```{r cleanNass}
# load data
outdir <- 'C:/Users/deinesji/Dropbox/1PhdJill/hpa/irrigation/data/nass'
outfile <- 'NASS_FieldCrops_County_RRB_raw_90_47counties.csv'
nass.df <- read.csv(paste0(outdir, '/', outfile), stringsAsFactors=F)

# # get a list of commodities for Anthony
#  # remove (D) values, remove commas (!), convert to numeric
#   nass.df <- nass.df[!grepl('(D)',nass.df$Value),]
#   nass.df$Value <- gsub(",", "", nass.df$Value, fixed = T)
#   nass.df$Value <- as.numeric(nass.df$Value)
#   
#   # keep only records with "ACRES HARVESTED"
#   phrase <- 'ACRES HARVESTED'
#   nass.df <- nass.df[grepl(phrase,nass.df$short_desc),]
# 
#   commodities <- unique(nass.df$commodity_desc)
#   outfile2 <- 'NASS_commodities_Anthony.csv'
#   write.csv(commodities, paste0(outdir, '/', outfile2), row.names=F)

# clean data
rrb2002 <- cleanNassCensusCounty(nass.df, year = 2002)
rrb2007 <- cleanNassCensusCounty(nass.df, year = 2007)
rrb2012 <- cleanNassCensusCounty(nass.df, year = 2012)

# get irrigated acreage by county
rrb2002.irr <- aggregate(IRRIGATED ~ fips5, data=rrb2002, FUN='sum')
rrb2007.irr <- aggregate(IRRIGATED ~ fips5, data=rrb2007, FUN='sum')
rrb2012.irr <- aggregate(IRRIGATED ~ fips5, data=rrb2012, FUN='sum')

# add years
rrb2002.irr$year <- 2002
rrb2007.irr$year <- 2007
rrb2012.irr$year <- 2012
```

## Compare!!

Here we go: compare irrigated areas by county

### Multipanel NASS

```{r multipanel_bigger_usda, fig.width = 8, fig.height = 3}
# merge NASS and RS data
countyAll.2002 <- countyAll[countyAll$year == 2002,]
merged2002 <- merge(rrb2002.irr, countyAll.2002[,c('fips5','irrArea.km2')])

countyAll.2007 <- countyAll[countyAll$year == 2007,]
merged2007 <- merge(rrb2007.irr, countyAll.2007[,c('fips5','irrArea.km2')])

countyAll.2012 <- countyAll[countyAll$year == 2012,]
merged2012 <- merge(rrb2012.irr, countyAll.2012[,c('fips5','irrArea.km2')])

allyears <- rbind(merged2002, merged2007, merged2012)

# plot
ggplot(allyears, aes(x=IRRIGATED,y=irrArea.km2)) +
  geom_point() +
  geom_abline(slope = 1, intercept = 0) + 
  stat_smooth_func(geom="text",method="lm",hjust=0,parse=TRUE, xpos = 100, ypos = 800) +
  #geom_smooth(method="lm",se=FALSE) +
  facet_wrap(~year, nrow=1) +
  ylab(expression(paste("HP-AIM Area (", km^2,")"))) + 
  xlab(expression(paste('USDA County Area (', km^2,')'))) + theme_bw() +
       theme(text = element_text(size = 15))  
```

### USGS Waterdata
just USGS data

```{r usgs, fig.height = 6}
# load usgs water data (compiled by Anthony)
usgsdir <- 'C:/Users/deinesji/Dropbox/1PhdJill/hpa/irrigation/data/nass/USGS_waterdata_county'
usgsname <- 'irrigation_water_use_by_county_1985_2010_w_acreage.csv'
usgsAll <- read.csv(paste0(usgsdir,'/',usgsname))

# make a fips column
usgsAll$fips5 <- paste0(sprintf("%02d",usgsAll$State_Code),
                        sprintf("%03d",usgsAll$Area_Code))

# subset to the 35 county fips
usgsIn <- usgsAll[usgsAll$fips5 %in% fips.in,]

# convert "thousand acres" to irrArea.km2
usgsIn$IRRIGATED <- usgsIn$IR.IrTot * 1000 * 0.00404686

# subset to relevant years
usgsYears <- usgsIn[usgsIn$Year %in% c(2000,2005,2010),]

# subset columns
usgs <- usgsYears[,c('fips5', 'IRRIGATED','Year')]
names(usgs)[3] <- 'year'

# merge with remote sensing data -------------
usgs$yearfips <- paste0(usgs$year, '-', usgs$fips5)

countyUSGS <- countyAll[countyAll$year %in% c(2000,2005,2010),]
countyUSGS$yearfips <- paste0(countyUSGS$year, '-', countyUSGS$fips5)

mergedUSGS <- merge(usgs, countyUSGS[,c('yearfips','irrArea.km2')])
mergedUSGS <- mergedUSGS[,c('fips5','IRRIGATED', 'year', 'irrArea.km2')]

# add source
mergedUSGS$source <- 'usgs'
```

plot usgs data vs remote sensing, by year 

```{r usgs_panel,fig.width = 8, fig.height = 3}
# plot
ggplot(mergedUSGS, aes(x=IRRIGATED,y=irrArea.km2)) +
  geom_point() +
  geom_abline(slope = 1, intercept = 0) + 
  stat_smooth_func(geom="text",method="lm",hjust=0,parse=TRUE, xpos = 100, ypos = 800) +
  #geom_smooth(method="lm",se=FALSE) +
  facet_wrap(~year, nrow=1) +
  ylab(expression(paste("HP-AIM Area (", km^2,")"))) + 
  xlab(expression(paste('USGS County Area (', km^2,')'))) + theme_bw() +
       theme(text = element_text(size = 15))  
```

multipanel with nass data

```{r nassusgs, fig.width = 8, fig.height = 6, dev=c('png','pdf')}
# combine with NASS data
allyears$source <- 'NASS'

countyValid <- rbind(allyears, mergedUSGS)
     
# plot
ggplot(countyValid, aes(x=IRRIGATED,y=irrArea.km2)) +
  geom_point() +
  geom_abline(slope = 1, intercept = 0) + 
  stat_smooth_func(geom="text",method="lm",hjust=0,parse=TRUE, xpos = 100, ypos = 800) +
  #geom_smooth(method="lm",se=FALSE) +
  facet_wrap(~year, nrow=2) +
  ylab(expression(paste("HP-AIM Area (", km^2,")"))) + 
  xlab(expression(paste('County Statistics (', km^2,')'))) + theme_bw() +
       theme(text = element_text(size = 15))
```

Change color of facet label strip by dataset. taken from http://stackoverflow.com/questions/19440069/ggplot2-facet-wrap-strip-color-based-on-variable-in-data-set

```{r nassusgs_colored, fig.width = 5, fig.height = 4}
library(gtable)
library(grid)
# store main plot
p1 <- ggplot(countyValid, aes(x=IRRIGATED,y=irrArea.km2)) +
  geom_point() +
  geom_abline(slope = 1, intercept = 0) + 
  stat_smooth_func(geom="text",method="lm",hjust=0,parse=TRUE, 
                   xpos = 100, ypos = 800) +
  #geom_smooth(method="lm",se=FALSE) +
  facet_wrap(~year, nrow=2) +
  ylab(expression(paste("HP-AIM Area (", km^2,")"))) + 
  xlab(expression(paste('County Statistics (', km^2,')'))) + theme_bw() +
       theme(text = element_text(size = 12))

# dummy plot
dummy <- ggplot(countyValid, aes(x=IRRIGATED,y=irrArea.km2)) +
  facet_wrap(~year, nrow=2) +
  geom_rect(aes(fill=source), xmin=-Inf, xmax = Inf, ymin = -Inf, ymax=Inf) +
  theme_minimal()

# gtable stuff
g1 <- ggplotGrob(p1)
g2 <- ggplotGrob(dummy)

gtable_select <- function (x, ...) 
{
  matches <- c(...)
  x$layout <- x$layout[matches, , drop = FALSE]
  x$grobs <- x$grobs[matches]
  x
}

panels <- grepl(pattern="panel", g2$layout$name)
strips <- grepl(pattern="strip-t", g2$layout$name)
g2$layout$t[panels] <- g2$layout$t[panels] - 1
g2$layout$b[panels] <- g2$layout$b[panels] - 1

new_strips <- gtable_select(g2, panels | strips)
grid.newpage()
grid.draw(new_strips)

gtable_stack <- function(g1, g2){
  g1$grobs <- c(g1$grobs, g2$grobs)
  g1$layout <- transform(g1$layout, z= z-max(z), name="g2")
  g1$layout <- rbind(g1$layout, g2$layout)
  g1
}
## ideally you'd remove the old strips, for now they're just covered
new_plot <- gtable_stack(g1, new_strips)
grid.newpage()
grid.draw(new_plot)

```

## error in space

```{r space, fig.height = 6, fig.width = 4}
# get county bounds
gisDir <- "C:/Users/deinesji/Dropbox/1PhdJill/hpa/irrigation/data/GIS/kml/shapefile"
cnty <- readOGR(gisDir, 'Counties_RRCAextent_Clip', verbose=F)

# calculate difference from 1:1 (rs - NASS)
merged2002$diff_2002 <- merged2002$irrArea.km2 - merged2002$IRRIGATED
hist(merged2002$diff_2002, main = '2002: RS - NASS')

merged2007$diff_2007 <- merged2007$irrArea.km2 - merged2007$IRRIGATED
hist(merged2007$diff_2007, main = '2007: RS - NASS')

merged2012$diff_2012 <- merged2012$irrArea.km2 - merged2012$IRRIGATED
hist(merged2012$diff_2012, main = '2012: RS - NASS')

# calculate percent error
merged2002$percErr_2002 <- (merged2002$irrArea.km2 - merged2002$IRRIGATED) /
                              merged2002$IRRIGATED * 100
hist(merged2002$percErr_2002, main = '2002: RS - NASS/NASS')

merged2007$percErr_2007 <- (merged2007$irrArea.km2 - merged2007$IRRIGATED) /
                              merged2007$IRRIGATED * 100
hist(merged2007$percErr_2007, main = '2007: RS - NASS/NASS')

merged2012$percErr_2012 <- (merged2012$irrArea.km2 - merged2012$IRRIGATED) /
                              merged2012$IRRIGATED * 100
hist(merged2012$percErr_2012, main = '2012: RS - NASS/NASS')

# add difference to spatial map
cntydiff02 <- merge(cnty, merged2002)
cntydiff0207 <- merge(cntydiff02, merged2007, by = 'fips5')
cntydiffall <- merge(cntydiff0207, merged2012, by = 'fips5')

# plot
palette <- brewer.pal(11, 'RdBu') 
palette2 <- c("#67001F", "#B2182B", "#D6604D", "#F4A582", "#FDDBC7",
              "gray90",
              "#D1E5F0", "#92C5DE", "#4393C3", "#2166AC", "#053061")

# difference
ticks = seq(-400,400,(400+400)/11)
spplot(cntydiffall, c('diff_2002','diff_2007','diff_2012'), 
       col.regions = palette2, 
       at = ticks, main = 'Difference: RS minus NASS')

ticks = seq(-200,200,(400)/11)
spplot(cntydiffall, c('percErr_2002','percErr_2007','percErr_2012'), 
       col.regions = palette2, 
       at = ticks, main = 'Percent Error (RS-NASS/NASS)')
```


