---
title: "NASS vs Remote Sensing Validation"
author: "Jill Deines"
date: "Friday, November 18, 2016"
output: 
  html_document:
    toc: yes
---

Goal: Compare irrigated acreages between remote sensing products and USDA NASS Ag Census Data. NASS data acquired through the Quick Stats API.


```{r knitrOpts, echo=FALSE}
library(knitr)
opts_chunk$set(cache=TRUE, cache.path='cache/2.1_NASS_v/',
               fig.path='figure/2.1_NASS_v/')
```

**R Packages Needed**

* `httr`: Vignette at https://cran.r-project.org/web/packages/httr/vignettes/quickstart.html

```{r packages, message=FALSE, echo=TRUE}
# for API interaction
library(httr) 
library(jsonlite)
# for county selection:
library(rgdal)
library(rgeos)
library(maps)
# for data manipulation
library(tidyr)
library(ggplot2)
```

## Functions: NASS Quick Stats API 
This script builds on work in 4.1_USDA_NASS.Rmd in the RRB_Rcode r project.

* API website: http://quickstats.nass.usda.gov/api
* My API key: 28EAA9E6-8060-34A4-981A-B2ED4692228A

Here, I'll make functions to ID counties of interest and pull NASS data from the API, and apply it over the counties.
