---
title: "Figures"
author: "Jill Deines"
date: "Friday, December 02, 2016"
output: 
  html_document:
    toc: yes
---

Goal: Make figures of irrigation frequency over the region


```{r knitrOpts, echo=FALSE}
library(knitr)
opts_chunk$set(cache=FALSE, cache.path='cache/4.1_FreqFigs/test5/',
               fig.path='figure/4.1_FreqFigs_randFor/to2016/test5/')
```

**R Packages Needed**

```{r packages, message=FALSE, echo=TRUE, eval=TRUE}
library(rgdal)
library(raster)
library(grid)
library(colorRamps)
library(fields)
library(RColorBrewer)
library(latticeExtra)
library(viridis)
library(rasterVis)
```

## Irrigation frequency
Exported raster of years irrigated from 06.0_summaryMaps in GEE.

### User variables

```{r userVars}
# dirs and names
rasDir <- 'C:/Users/deinesji/Google Drive/GEE_classification/RRB_test5_randFor_interannualClean'
rasRaw <- 'IrrigationFrequencyMap_0and1neg.tif'
rasMasked <- 'IrrigationFrequencyMap_0and1neg_boundMask.tif'
  
# inset sets
rasSmall1.name <- 'IrrigationFrequencyMap_0and1neg_inset.tif'
earlyInset1.name <- 'earliestYearIrrigated_inset.tif'
lateInset1.name <- 'latestYearIrrigated_inset.tif'

# rasSmall2.name <- 'IrrigationFrequencyMap_0and1neg_inset_2.tif'
# earlyInset2.name <- 'earliestYearIrrigated_inset_2.tif'
# lateInset2.name <- 'latestYearIrrigated_inset_2.tif'
# 
# rasSmall3.name <- 'IrrigationFrequencyMap_0and1neg_inset_3.tif'
# earlyInset3.name <- 'earliestYearIrrigated_inset_3.tif'
# lateInset3.name <- 'latestYearIrrigated_inset_3.tif'

# extent shapefile
extentDir <- 'C:/Users/deinesji/Dropbox/1PhdJill/hpa/irrigation/data/GIS/kml/shapefile'
extentName <- 'BigExtent_RRB_RRCA_buff'
```

Set some common color scales, etc, among figures

```{r commonPlotVars}
# number of years color ramps
numYears <- 18
#backgroundcolor <- 'gray60'
backgroundcolor <- 'gray32'

# number of years color ramp
matlab <- matlab.like2(numYears+1)
#matlab <- viridis(numYears+1)

# year color ramp
# red yellow blue for dave
yearramp <- rev(colorRampPalette(brewer.pal(9,'RdYlBu'))(numYears))

##green/pink for me
#yearramp <- rev(colorRampPalette(brewer.pal(9,'PiYG'))(numYears))
```

end user vars

### map data pre-processing - only need to run 1x
First, load raster domain raster and mask, and write out masked version to save time in the future.

```{r frequencyPrep, eval=FALSE}
# load full raster 
freq <- raster(paste0(rasDir,'/',rasRaw))

# load extent 
extent0 <- readOGR(extentDir, extentName, verbose=F)
extent <- spTransform(extent, CRS(proj4string(freq)))

# set -1 to 0 (areas masked by nlcd)
freq[freq == -1] <- 0

# mask by polygon bound
freq2 <- mask(freq, extent)

# write out masked raster
writeRaster(freq2, paste0(rasDir,'/',rasMasked))
```

### Plots

### histogram
Frequency histogram of number of years irrigated for pixels: get frequency of values

```{r histPrep, eval=FALSE}
# load full raster
freqHist <- raster(paste0(rasDir,'/',rasRaw))
# count by classes
yearCounts <- freq(freqHist)
# takes a while, save output
write.csv(yearCounts, paste0(rasDir,'/test5_randFor_yearMapCounts.csv'),
          row.names=F) 
```

plot it

```{r hist, fig.width = 6, fig.height = 5, eval=TRUE}
# load yearmap counts
yearCounts <- read.csv(paste0(rasDir,'/test5_randFor_yearMapCounts.csv'))
# sort
yearCounts <- yearCounts[order(yearCounts$value),]

# remove 0 and -1 row
yearCounts2 <- yearCounts[!(yearCounts$value == -1 | yearCounts$value == 0),]
rownames(yearCounts2) <- 1:nrow(yearCounts2)


# get a percentage column: of area ever irrigated
yearCounts2$percentOfIrr <- yearCounts2$count/sum(yearCounts2$count) * 100

# side note: what percentage for 16-18 years? (perennial)
sum(yearCounts2$count[15:17])/sum(yearCounts2$count)

# non-perennial
sum(yearCounts2$count[1:14])/sum(yearCounts2$count)

# arange plot parameters
bardata <- yearCounts2$percentOfIrr
colors <- matlab[2:19]

# vert plot
barplot(bardata, col=colors, space=0, las=1, line=-.5)
mtext(side = 2, "Percent of Irrigated Area", line=2, cex=1.1)
mtext(side = 1, "Years Irrigated", line=4, cex=1.1)

# legend
image.plot(z=0:18,add=F,legend.only=T,  col=c(backgroundcolor,matlab[1:18]), 
           horizontal = T,
          smallplot=c(0.08,.902,.13,.18), 
         axis.args=list(at=c(0,3,6,9,12,15,18), 
                          labels=as.character(c(0,3,6,9,12,15,18))))

```

hist without colorscale

```{r hist_noScale, fig.width = 6.5, fig.height = 5, eval=TRUE}

# vert plot
barplot(bardata, col=colors, space=0.1, las=1, line=-.5,
        names.arg=c('2','','4','','6','','8','','10','','12',
                    '','14','','16','','18'))
mtext(side = 2, "Percent of Irrigated Area", line=2, cex=1.1)
mtext(side = 1, "Years Irrigated", line=2, cex=1.1)

```


### Number of Years Irrigated Maps

Number of years irrigated figure

```{r regionLatLong,fig.width = 10, fig.height = 6.5}
extent0 <- readOGR(extentDir, extentName, verbose=F)
extent.ll  <- spTransform(extent0, CRS("+proj=longlat +datum=WGS84"))
plot(extent.ll, axes=T, las=1)       
#plot(freq2)
```


```{r freqMap_matlab2_bottom_matlab, fig.width = 10, fig.height = 5}
# load full raster
freq2 <- raster(paste0(rasDir,'/',rasMasked))

# plot scale
scale = list("SpatialPolygonsRescale", layout.scale.bar(), 
  offset = c(-225000,1760000), scale = 100000, fill=c("transparent","black"))
text1 = list("sp.text", c(-220000,1750000), "0", cex=1.2)
text2 = list("sp.text", c(-125000,1750000), "100 km",cex = 1.2)
arrow = list("SpatialPolygonsRescale", layout.north.arrow(type=1), offset = c(-223000, 
    1780000), scale = 40000)

# color ramp 
spplot(freq2, col.regions = c(backgroundcolor,matlab), maxpixels = 2000000,
       sp.layout=list(scale, text1, text2, arrow), 
       par.settings=list(axis.line=list(col=NA)), 
       #scales = list(draw=T),
       colorkey=F)#list(labels=list(cex=1.2),
                  #at = c(0:18))

levelplot(freq2, col.regions=c(backgroundcolor,matlab))

image.plot(z=0:18,add=F,legend.only=T,  col=c(backgroundcolor,matlab), 
           horizontal = T,
          smallplot=c(0.05,.95,.15,.2), 
         axis.args=list(at=c(0,3,6,9,12,15,18), 
                          labels=as.character(c(0,3,6,9,12,15,18))))


# scale bar legend title
#grid.text('Years Irrigated',x=unit(0.975, "npc"),y=unit(0.5, 'npc'), rot=-90, gp=gpar(fontsize=18))
```


Integrated colorbar on right (retains map frame border)

```{r freqMap_matlab2_right, fig.width = 10, fig.height = 5, dpi=96, eval=FALSE}
# plot scale
scale = list("SpatialPolygonsRescale", layout.scale.bar(), 
  offset = c(-225000,1760000), scale = 100000, fill=c("transparent","black"))
text1 = list("sp.text", c(-220000,1750000), "0")
text2 = list("sp.text", c(-125000,1750000), "100 km")


spplot(freq, col.regions = c(backgroundcolor,matlab), maxpixels = 2000000,
       par.settings = list(layout.widths=list(right.padding=4.5)),
       sp.layout=list(scale, text1, text2),
       colorkey=list(labels=list(cex=1.2),
                     at = c(0:18)))


# scale bar legend title
grid.text('Years Irrigated',x=unit(0.975, "npc"),y=unit(0.5, 'npc'), rot=-90, gp=gpar(fontsize=18))
```

#### inset: Years Irrigated Maps

3 location options

```{r inset1_freqMap_matlab, fig.width = 2.5, fig.height = 3.2, dpi=96, cache=FALSE}
# load full raster
freq_inset1 <- raster(paste0(rasDir,'/',rasSmall1.name))
freq_inset1[freq_inset1 == -1] <- 0



# manually make a scale polygon... :(
xmin <- -406500
ymin <- 1983400
scale <- 5000
height <- 1500

text1 = list("sp.text", c(xmin+400,ymin+300), "3 km", cex=.5)

scalePoly <- Polygon(coords = as.matrix(data.frame(
  lon = c(xmin, xmin, xmin+scale, xmin+scale, xmin),
  lat = c(ymin, ymin+height, ymin+height, ymin, ymin))),
  hole=F)
scalePoly2 <- Polygons(list(scalePoly), ID='scale')
scale.sp <- SpatialPolygons(list(scalePoly2), proj4string = CRS(proj4string(freq_inset1)))


# color ramp 
spplot(freq_inset1, col.regions = c(backgroundcolor,matlab), maxpixels = 200000,
       par.settings = list(layout.widths=list(right.padding=4.5)),
      # sp.layout=list(text1),
       colorkey=list(space = 'bottom',
                     cex = 1.2,
                     at=c(0:18), 
                     labels=c('0','','','3','','','6','','','9',
                              '','','12','','','15','','','18'),
                     width = .85,
                     tck = .7)) +
  latticeExtra::layer(sp.polygons(scale.sp, fill='white', col='black'))
```


```{r inset2_freqMap, fig.width = 2.5, fig.height = 3.2, dpi=96, cache=FALSE, eval=FALSE}
# load full raster
freq_inset2 <- raster(paste0(rasDir,'/',rasSmall2.name))
freq_inset2[freq_inset2 == -1] <- 0

# color ramp 
spplot(freq_inset2, col.regions = c(backgroundcolor,matlab), maxpixels = 200000,
       par.settings = list(layout.widths=list(right.padding=4.5)),
       #sp.layout=list(scale, text1, text2),
       colorkey=list(space = 'bottom',
                     cex = 1.2,
                     at=c(0:18), 
                     labels=c('0','','','3','','','6','','','9',
                              '','','12','','','15','','','18'),
                     width = .85,
                     tck = .7))
```

```{r inset3_freqMap, eval=FALSE, fig.width = 2.5, fig.height = 3.2, dpi=96, cache=FALSE}
# load full raster
freq_inset3 <- raster(paste0(rasDir,'/',rasSmall3.name))
freq_inset3[freq_inset3 == -1] <- 0

# color ramp 
spplot(freq_inset3, col.regions = c(backgroundcolor,matlab), maxpixels = 200000,
       par.settings = list(layout.widths=list(right.padding=4.5)),
       #sp.layout=list(scale, text1, text2),
       colorkey=list(space = 'bottom',
                     cex = 1.2,
                     at=c(0:18), 
                     labels=c('0','','','3','','','6','','','9',
                              '','','12','','','15','','','18'),
                     width = .85,
                     tck = .7))
```

#### inset: Earliest Year Irrigated maps


```{r inset1_early, fig.width = 2.5, fig.height = 3.2, dpi=96, cache=FALSE}
# load full raster
early_inset1 <- raster(paste0(rasDir,'/',earlyInset1.name))
early_inset1[early_inset1 == 0] <- NA

# color ramp 
spplot(early_inset1, col.regions = c(yearramp), maxpixels = 200000,
       par.settings = list(layout.widths=list(right.padding=4.5),
                           panel.background=list(col=backgroundcolor)),
       #sp.layout=list(scale, text1, text2),
       colorkey=list(space = 'bottom',
                     cex = 1.2,
                     at=c(1999:2016),
                     labels=c('','2000','','','','','2005','','','',
                              '','2010','','','','','2015',''),
                     width = .85,
                     tck = .7)) #+
  #layer(sp.polygons(extent, fill=backgroundcolor), under=T)
```

```{r inset2_early, fig.width = 2.5, fig.height = 3.2, dpi=96, cache=FALSE, eval=FALSE}
# load full raster
early_inset2 <- raster(paste0(rasDir,'/',earlyInset2.name))
early_inset2[early_inset2 == 0] <- NA

# color ramp 
spplot(early_inset2, col.regions = c(yearramp), maxpixels = 200000,
       par.settings = list(layout.widths=list(right.padding=4.5),
                           panel.background=list(col=backgroundcolor)),
            colorkey=list(space = 'bottom',
                     cex = 1.2,
                     at=c(1999:2016),
                     labels=c('','2000','','','','','2005','','','',
                              '','2010','','','','','2015',''),
                     width = .85,
                     tck = .7)) #+
```

```{r inset3_early, fig.width = 2.5, fig.height = 3.2, dpi=96, cache=FALSE, eval=FALSE}
# load full raster
early_inset3 <- raster(paste0(rasDir,'/',earlyInset3.name))
early_inset3[early_inset3 == 0] <- NA

# color ramp 
spplot(early_inset3, col.regions = c(yearramp), maxpixels = 200000,
       par.settings = list(layout.widths=list(right.padding=4.5),
                           panel.background=list(col=backgroundcolor)),
             colorkey=list(space = 'bottom',
                     cex = 1.2,
                     at=c(1999:2016),
                     labels=c('','2000','','','','','2005','','','',
                              '','2010','','','','','2015',''),
                     width = .85,
                     tck = .7)) #+
```

#### inset: Latest Year Irrigated maps

3 location options

```{r inset1_late, fig.width = 2.5, fig.height = 3.2, dpi=96, cache=FALSE}
# load full raster
late_inset1 <- raster(paste0(rasDir,'/',lateInset1.name))
late_inset1[late_inset1 == 0] <- NA

# color ramp 
spplot(late_inset1, col.regions = c(yearramp), maxpixels = 200000,
       par.settings = list(layout.widths=list(right.padding=4.5),
                           panel.background=list(col=backgroundcolor)),
             colorkey=list(space = 'bottom',
                     cex = 1.2,
                     at=c(1999:2016),
                     labels=c('','2000','','','','','2005','','','',
                              '','2010','','','','','2015',''),
                     width = .85,
                     tck = .7)) #+
```

```{r inset2_late, fig.width = 2.5, fig.height = 3.2, dpi=96, cache=FALSE, eval=FALSE}
# load full raster
late_inset2 <- raster(paste0(rasDir,'/',lateInset2.name))
late_inset2[late_inset2 == 0] <- NA

# color ramp 
spplot(late_inset2,col.regions = c(yearramp), maxpixels = 200000,
       par.settings = list(layout.widths=list(right.padding=4.5),
                           panel.background=list(col=backgroundcolor)),
              colorkey=list(space = 'bottom',
                     cex = 1.2,
                     at=c(1999:2016),
                     labels=c('','2000','','','','','2005','','','',
                              '','2010','','','','','2015',''),
                     width = .85,
                     tck = .7)) #+
```

```{r inset3_late, fig.width = 2.5, fig.height = 3.2, dpi=96, cache=FALSE, eval=FALSE}
# load full raster
late_inset3 <- raster(paste0(rasDir,'/',lateInset3.name))
late_inset3[late_inset3 == 0] <- NA

# color ramp 
spplot(late_inset3, col.regions = c(yearramp), maxpixels = 200000,
       par.settings = list(layout.widths=list(right.padding=4.5),
                           panel.background=list(col=backgroundcolor)),
             colorkey=list(space = 'bottom',
                     cex = 1.2,
                     at=c(1999:2016),
                     labels=c('','2000','','','','','2005','','','',
                              '','2010','','','','','2015',''),
                     width = .85,
                     tck = .7)) #+
```





### Stand alone color ramps

```{r colorbar_matlab,fig.width = 8, fig.height = 4}
colRamp <- c(backgroundcolor,matlab.like2(19))

plot(1:10)
image.plot(z=0:18,add=F,legend.only=T, nlevel=16,  col=colRamp, 
           horizontal = T,
           smallplot=c(0.1,.9,.93,.98), 
         axis.args=list(at=c(0,3,6,9,12,15,18), 
                          labels=as.character(c(0,3,6,9,12,15,18))))
```


```{r colorbar_PiYG,fig.width = 5, fig.height = 4}
colRamp <- rev(colorRampPalette(brewer.pal(11,'PiYG'))(14))

plot(1:10)
image.plot(z=0:13,add=F,legend.only=T, nlevel=14,  col=colRamp, 
           horizontal = T,
           smallplot=c(0.1,.9,.93,.98), 
           axis.args=list(at=c(1,3,5,7,9,11,13), 
                          labels=as.character(c(2000,2002,2004,2006,2008,2010,2012))))
```

```{r colorbar_PRGn,fig.width = 5, fig.height = 4}
colRamp <- rev(colorRampPalette(brewer.pal(11,'PRGn'))(14))

plot(1:10)
image.plot(z=0:13,add=F,legend.only=T, nlevel=14,  col=colRamp, 
           horizontal = T,
           smallplot=c(0.1,.9,.93,.98), 
           axis.args=list(at=c(1,3,5,7,9,11,13), 
                          labels=as.character(c(2000,2002,2004,2006,2008,2010,2012))))
```
