---
title: "Figure: Years Irrigated"
author: "Jill Deines"
date: "March 16, 2017"
output: 
  html_document:
    toc: yes
---

Goal: Make maps of irrigation frequency over the region

Use: 

* update figure path (make a MI folder; line 21) and user variables
* run "mapPrep" chunk within R (like normal)
* hit 'knit html' to run documents; figures will be saved in figure folder
* i get an error when knitting but the figures come out fine so ::shrug::

Note: it takes a while to output full resolution figures. If you want to run it quickly just to get a vague idea about things, run it with a low 'maxpixel' user variable.


```{r knitrOpts, echo=FALSE}
library(knitr)
opts_chunk$set(cache=FALSE, dev = 'png',
               fig.path='S:/Users/xutianfa/Irrigation/poster_files/figures/RRB/YearMap/')
```

**R Packages Needed**

```{r packages, message=FALSE, echo=TRUE, eval=TRUE}
library(rgdal)
library(raster)
library(grid)
library(colorRamps)
library(fields)
library(RColorBrewer)
library(rgdal)
library(latticeExtra)
```

## User variables
If changing the full map width and height (maybe necessary for MI)...you may need to adjust the color bar positioning in the map with legend. That's an irritating, iterative process. 

```{r userVars}
# raster data directory
rasDir <- 'S:/Users/xutianfa/Irrigation/poster_files/data/RRB'
# name of exported raster for full domain
rasAll <- 'IrrigationFrequencyMap_0and1neg.tif'
# name of inset raster
rasInset <- 'IrrigationFrequencyMap_0and1neg_inset.tif'

# name of raster with masked domain
rasMask <- 'IrrigationFrequencyMap_0and1neg_boundMask.tif'

# extent shapefile
shapeDir <- 'S:/Users/deinesji/HPA/gis'
shapename <- 'BigExtent_RRB_RRCA_buff'


# figure dimensions and options
fullMapWidth = 10
fullMapHeight = 5
insetWidth = 4
insetHeight = 4
dpii = 300
maxpixel = 1500000 # defines map resolution
backgroundcolor <- 'gray32'
```


## Data Prep
We tallied number of years each pixel was irrigated in GEE using the script in the MSU_Hydrogeology/FigureScripts shared repository.

First, we load the raster and set pixels outside of the model domain (value = -1) to NA, which for dumb reasons can't be done in GEE. Then we write this version out for easy use later, since masking in R takes a while.

Just run this once within R normally; chunk is set to 'false' so it doesn't get run every time you knit the document to produce figures

```{r mapPrep, eval=FALSE}
# load full raster and set -1 to NA (basically mask areas outside of Extent)
freq <- raster(paste0(rasDir,'/',rasAll))
freq[freq == -1] <- NA
# write out masked raster
writeRaster(freq, paste0(rasDir,'/', rasMask))
```


## Number of Years Irrigated Maps

vertical legend, frame

```{r yearMap_right, fig.width = fullMapWidth, fig.height = fullMapHeight, dpi=dpii}
# load full raster
freq <- raster(paste0(rasDir,'/',rasMask))

# load shapefile
spdf <- readOGR(shapeDir, shapename, verbose=F)
spdf <- spTransform(spdf, CRS(proj4string(freq)))

# plot scale
scale = list("SpatialPolygonsRescale", layout.scale.bar(), 
  offset = c(-225000,1760000), scale = 100000, fill=c("transparent","black"))
text1 = list("sp.text", c(-220000,1745000), "0")
text2 = list("sp.text", c(-125000,1745000), "100 km")

# color ramp 
matlab <- matlab.like2(19)

# plot
spplot(freq, col.regions = c(backgroundcolor,matlab), maxpixels = maxpixel,
       par.settings = list(layout.widths=list(right.padding=4.5)),
       sp.layout=list(scale, text1, text2),
       colorkey=list(labels=list(cex=1.2),
                     at = c(0:18)))# +
#layer(sp.polygons(spdf, fill='transparent', col='black', alpha=0.7))

# scale bar legend title
grid.text('Years Irrigated',x=unit(0.975, "npc"),y=unit(0.5, 'npc'), rot=-90, gp=gpar(fontsize=18))
```

No frame, seperate horizontal legend

```{r yearMap_bottom, fig.width = fullMapWidth, fig.height = fullMapHeight, dpi=dpii}
# load full raster
freq <- raster(paste0(rasDir,'/',rasMask))

# plot scale
scale = list("SpatialPolygonsRescale", layout.scale.bar(), 
  offset = c(-225000,1760000), scale = 100000, fill=c("transparent","black"))
text1 = list("sp.text", c(-220000,1750000), "0")
text2 = list("sp.text", c(-125000,1750000), "100 km")

# color ramp 
matlab <- matlab.like2(19)

# plot
spplot(freq, col.regions = c(backgroundcolor,matlab), maxpixels = maxpixel,
       sp.layout=list(scale, text1, text2), 
       par.settings=list(axis.line=list(col=NA)), 
       colorkey=F) #+
#layer(sp.polygons(spdf, fill='transparent', col='black', alpha=0.7))

image.plot(z=0:18,add=F,legend.only=T,  col=c(backgroundcolor,matlab), 
           horizontal = T,
          smallplot=c(0.05,.95,.15,.2), 
         axis.args=list(at=c(0,3,6,9,12,15,18), 
                          labels=as.character(c(0,3,6,9,12,15,18))))

```

### inset

```{r yearMap_inset, fig.width = insetWidth, fig.height = insetHeight, dpi=dpii}
# load full raster
freq_inset <- raster(paste0(rasDir,'/',rasInset))

# # plot scale
# scale = list("SpatialPolygonsRescale", layout.scale.bar(), 
#   offset = c(-225000,1760000), scale = 100000, fill=c("transparent","black"))
# text1 = list("sp.text", c(-220000,1750000), "0")
# text2 = list("sp.text", c(-125000,1750000), "100 km")

# color ramp 
spplot(freq_inset, col.regions = c(backgroundcolor,matlab), maxpixels = 200000,
       par.settings = list(layout.widths=list(right.padding=4.5)),
       #sp.layout=list(scale, text1, text2),
       colorkey=F)

```


