---
title: "Figures Irrigated volume v area"
output: 
  html_document:
    toc: yes
---

Goal: plot times series of irrigated areas vs pumping volumes

update: 4/9/17 - main purpose now is to create the RRCA pumping and area derived datasets


```{r knitrOpts, echo=FALSE}
library(knitr)
opts_chunk$set(cache=FALSE, cache.path='cache/4.21_VolumeAreaFigs_rf/',
               fig.path='figure/4.21_VolumeAreaFreqFigs_rf/')
```

**R Packages Needed**

```{r packages, message=FALSE, echo=TRUE, eval=TRUE}
library(rgdal)
library(raster)
library(rgeos)
library(grid)
library(colorRamps)
library(ggplot2)
library(tidyr)
library(dplyr)
source('functions/stat_smooth_func_all.R')
```

## load and process data

Pumping data from the RRB_Rcode project (4.3_Pumping_Data.Rmd). This file is data for the Republican Basin, by state.

UPDATE 2/20/17: big change to use combined gw/sw dataset

UPDATE 10/17/17: checking fixed irrigation water data

```{r loadPumpingData_fixed}
# years wanted
startYear = 1999
endYear = 2015 # current dataset extent

# # pumping data extracted from the RRCA groundwater model folks, see readme notes
# pumpDir <- 'C:/Users/deinesji/Dropbox/1PhdJill/hpa/irrigation/data/pumping'
# #allIrr0 <- read.csv(paste0(pumpDir,'/RRCA_irrAllByCell_1918-2015_longFormat_gw_sw_co.csv'))
# allIrr0 <- read.csv(paste0(pumpDir,'/RRCA_irrAllByCell_1918-2015_longFormat_gw_sw_co_FIXED_20171017.csv'))
# 
# 
# # subset years
# allIrr <- allIrr0[allIrr0$Year >= startYear & allIrr0$Year <= endYear,]
# 
# # write out subset for future easier use
# write.csv(allIrr, row.names=F,
#           paste0(pumpDir,'/RRCA_irrAllByCell_1999-2015_longFormat_gw_sw_co_FIXED_20171017.csv'))

#load year-restricted dataset
pumpDir <- 'C:/Users/deinesji/Dropbox/1PhdJill/hpa/irrigation/data/pumping'
#allIrr <- read.csv(paste0(pumpDir,'/RRCA_irrAllByCell_1999-2015_longFormat_gw_sw_co.csv'))
allIrr <- read.csv(paste0(pumpDir,'/RRCA_irrAllByCell_1999-2015_longFormat_gw_sw_co_FIXED_20171017.csv'))


# replace NA's with zeros
#allIrr[,18:23][is.na(allIrr[,18:23])] <- 0 
colsToZero <- c('swirr_m3','swirr_km3','swirr_acreft','coirr_m3','coirr_km3','coirr_acreft')
allIrr[,colsToZero][is.na(allIrr[,colsToZero])] <- 0 

# sum total water use
allIrr$total_km3 <- allIrr$gwirr_km3 + allIrr$swirr_km3 + allIrr$coirr_km3
allIrr$total_m3 <- allIrr$gwirr_m3 + allIrr$swirr_m3 + allIrr$coirr_m3

### sum areas by state RRCA --------------------------------------
## dplyr exmple
# irrStates.RRCA.km3 <- allIrr %>%
#                     group_by(state, year) %>%
#                     summarise(total = sum(total_km3),
#                               gw = sum(gwirr_km3),
#                               sw = sum(swirr_km3),
#                               co = sum(coirr_km3))

allIrr <- allIrr %>% rename(year = Year)
irrStates.RRCA.km3 <- aggregate(cbind(total_km3, gwirr_km3, swirr_km3, coirr_km3)
                            ~ state + year, data = allIrr, 
                           FUN = 'sum')

# add a year-state key to join data with irrigated area
irrStates.RRCA.km3$key <- paste0(irrStates.RRCA.km3$state, irrStates.RRCA.km3$year)

# plot for fun
ggplot(data = irrStates.RRCA.km3,
       aes(x = year, y = total_km3, colour=state)) +
  geom_line() + theme_bw() +
  ggtitle('Irrigation Volume by State RRCA')

### sum areas by state RRB basin --------------------------------------
# extract only RRB cells
irrRRB <- allIrr[!is.na(allIrr$inRRB),]

irrStates.RRB.km3 <- aggregate(cbind(total_km3, gwirr_km3, swirr_km3, coirr_km3)
                            ~ state + year, data = irrRRB, 
                           FUN = 'sum')

# add a year-state key to join data with irrigated area
irrStates.RRB.km3$key <- paste0(irrStates.RRB.km3$state, irrStates.RRB.km3$year)

# plot for fun
ggplot(data = irrStates.RRB.km3,
       aes(x = year, y = total_km3, colour=state)) +
  geom_line() + theme_bw() +
  ggtitle('Total Irrigation by State RRB')

# total pumping by type for full RRCA -----------------------------------
# sum pumping types by year
allIrr.years.km3 <- aggregate(cbind(total_km3, gwirr_km3, swirr_km3, coirr_km3)
                              ~ year, data = allIrr, FUN = 'sum')

# longify for ggplot
allIrr.years.long <- gather(allIrr.years.km3, key = 'type', value = 'km3',
                            gwirr_km3:coirr_km3)
allIrr.years.long$type2 <- substr(allIrr.years.long$type, 1,2)

# reorder to stack gw on bottom
changer <- data.frame(old = c('co','sw','gw'),
                      source = c('2_co','3_sw','1_gw'))
allIrr.years.long <- merge(allIrr.years.long, changer, by.x = 'type2',
                           by.y = 'old')

# plot stacked area
ggplot(data=allIrr.years.long,
       aes(x=year, y = km3, fill = source)) +
  geom_area(alpha = .7) +
  scale_fill_brewer(palette='Paired') +
  theme_bw() +
  ggtitle('Total Irrigation by Type, RRCA')

# export the cleaned datasets that retains state and type
derivedDir <- 'C:/Users/deinesji/Dropbox/1PhdJill/hpa/irrigation/data/derivedDatasets'
# write.csv(irrStates.RRCA.km3, row.names=F,
#          paste0(derivedDir, '/RRCA_fullModel_irrigation_volume_FIXED_20171017.csv'))
# write.csv(irrStates.RRB.km3, row.names=F,
#          paste0(derivedDir, '/RRB_byState_irrigation_volume_FIXED_20171017.csv'))
```

Make a 'master data' file for the RRCA boundary

```{r loadAreaData, eval=FALSE}
cellArea <- 30*30 # m^2

regionsDir <- 'C:/Users/deinesji/Google Drive/GEE_tableExports/RRB_test5_regional_county_Stats'

# get filenames
files <- list.files(regionsDir, pattern="*_test5_randFor_interannual_plusAncillary.csv")

# read in to a single dataframe
all1 = as.data.frame(do.call(rbind, lapply(files, function(x) {
              csv1 <- read.csv(paste0(regionsDir,'/',x))
              csv1$year <- as.numeric(substr(x, start=1,stop=4))
              return(csv1)
            })))

# convert cell count to area
all1$irrigated_m2 <- all1$X1 * cellArea
all1$irrigated_km2 <- all1$irrigated_m2 / 1000000

# state RRCA  data only
areaRRCA.states <- all1[grepl('_rrcaBound', all1$masterID),]

# keep fewer columns (location, year, irrigated area, annual (Dec-Aug) precip)
areaRRCA.states <- areaRRCA.states[,c('masterID','year','irrigated_km2','pr_ann')]

# add total area of RRCA region
rrca.spdf <- readOGR('C:/Users/deinesji/Dropbox/1PhdJill/hpa/irrigation/data/GIS/kml/RRCA_bound_states.kml', 
                     'layer', verbose = T)
rrca.spdf <- spTransform(rrca.spdf, CRS = '+init=epsg:5070')
rrca.spdf$area_m2 <- gArea(rrca.spdf, byid=T)
rrca.spdf$area_km2 <- rrca.spdf$area_m2 * 1e-6

# 3 values will be recycled, confirmed they are in the correct order
areaRRCA.states$totalArea_km2 <- rrca.spdf$area_km2


# add a year column and an id column to merge with pumping
areaRRCA.states$state <- substr(areaRRCA.states$masterID, start=1, stop=2)
areaRRCA.states$key <- paste0(areaRRCA.states$state,areaRRCA.states$year)

# export to derived dataset folder
derivedDir <- 'C:/Users/deinesji/Dropbox/1PhdJill/hpa/irrigation/data/derivedDatasets'
write.csv(areaRRCA.states, paste0(derivedDir, '/RRCA_fullModel_area-rfTest5_ppt-dec-aug.csv'), row.names=F)
```

4/9/2017 - end test 4 edits
4/26/2017 - end test 5 edits

## RRCA groundwater model boundary analyses
started 3/1/2017

```{r rrcaPlots, eval = FALSE}
# merge pumping and area data
plotData2 <- merge(areaRRCA.states, irrStates.RRCA.km3, by='key', all.x = T)
plotData2 <- plotData2[,c('year.x','state.x','irrigated_km2','total_km3','pr_ann')]

# format for ggplot
plotDataLong2 <- gather(plotData2, key = variable, value = value,
                       irrigated_km2:pr_ann)
```

---- end 3/1 edits - still needs updated for RRCA model bounds

## Republican Basin Boundary analyses

```{r stateBasinPlots, eval=FALSE}
# state basin data only
areaStates <- all1[grepl('basin', all1$masterID) & !grepl('NRD', all1$masterID),]

# keep fewer columns (location, year, irrigated area, annual (Dec-Aug) precip)
areaStates <- areaStates[,c('masterID','year','irrigated_km2','pr_ann')]

# add a year column and an id column to merge with pumping
areaStates$state <- substr(areaStates$masterID, start=1, stop=2)
areaStates$key <- paste0(areaStates$state,areaStates$year)

# merge pumping and area data
plotData <- merge(areaStates, irrStates.RRB.km3, by='key', all.x = T)
plotData <- plotData[,c('year.x','state.x','irrigated_km2','total_km3','pr_ann')]

# format for ggplot
plotDataLong <- gather(plotData, key = variable, value = value,
                       irrigated_km2:pr_ann)

# plot times series of pumping, area, and precip
state <- 'NE'

# check a linear model of irrigated area to precip
yarg = lm(irrigated_km2 ~ pr_ann, 
          data = plotData[plotData$state.x == state,])
summary(yarg)



ggplot(data = plotDataLong[plotDataLong$state.x == state,],
      aes(x = year.x, y = value)) +
  geom_line() +
  facet_wrap(~variable, nrow=3, scales='free_y') +
  theme_bw() + ggtitle('Nebraska Republican River Basin')

# plot times series of pumping, area, and precip
state <- 'KS'

ggplot(data = plotDataLong[plotDataLong$state.x == state,],
      aes(x = year.x, y = value)) +
  geom_line() +
  facet_wrap(~variable, nrow=3, scales='free_y') +
  theme_bw() + ggtitle('Kansas Republican River Basin')

# plot times series of pumping, area, and precip
state <- 'CO'

ggplot(data = plotDataLong[plotDataLong$state.x == state,],
      aes(x = year.x, y = value)) +
  geom_line() +
  facet_wrap(~variable, nrow=3, scales='free_y') +
  theme_bw() + ggtitle('Colorado Republican River Basin')

```

```{r fullRRB_plots, fig.height = 5.5, fig.width = 6, eval=FALSE}
# replace volume NA's with 0 for now
plotData[is.na(plotData$total_km3),'total_km3'] <- 0

# get irrigated area and pumping for full basin
plotDataRRB <- aggregate(cbind(irrigated_km2,total_km3) ~  year.x, 
                         data = plotData, FUN = 'sum', na.rm = T)
# and do the mean for precip
rrbPrecip <- aggregate(pr_ann ~  year.x, data = plotData, FUN = 'mean')


# combine and gather
plotDataRRB <- cbind(plotDataRRB, rrbPrecip[,'pr_ann'])
names(plotDataRRB) <- c('year','Irrigated_Area', 'Pumping', 'zPrecipitation')

# replace NA's for 2015 and 2016
plotDataRRB[plotDataRRB$Pumping == 0,'Pumping'] <- NA

# area over time lm
area.lm <- lm(Irrigated_Area ~ year, data = plotDataRRB)
summary(area.lm)

# pumping over time lm
pump.lm <- lm(Pumping ~ year, data = plotDataRRB)
summary(pump.lm)

# gather
plotDataRRBLong <- gather(plotDataRRB, key = variable, value = value,
                          Irrigated_Area:zPrecipitation)

# plots with no facet labels
ggplot(data = plotDataRRBLong,
      aes(x = year, y = value, colour=variable)) +
  geom_line() +
  scale_colour_manual(values=c('black','red','blue')) + 
  facet_wrap(~variable, nrow=3, scales='free_y') +
  scale_x_continuous(name = '', 
                     breaks = c(2000,2002,2004,2006,2008,2010,2012), 
                     labels = c(2000,2002,2004,2006,2008,2010,2012)) +
  ylab('') +
  theme_bw() + theme(legend.position = 'none',
                     strip.background=element_blank(),
                     strip.text.x = element_blank(),
                     text = element_text(size = 16))

# plots with facet labels
ggplot(data = plotDataRRBLong,
      aes(x = year, y = value, colour=variable)) +
  geom_line() +
  scale_colour_manual(values=c('black','red','blue')) + 
  facet_wrap(~variable, nrow=3, scales='free_y') +
    scale_x_continuous(name = '', 
                     breaks = c(2000,2002,2004,2006,2008,2010,2012), 
                     labels = c(2000,2002,2004,2006,2008,2010,2012)) +
  ylab('') +
  ylab('') +
  theme_bw() + theme(legend.position = 'none',
                     text = element_text(size = 16))
```

```{r fullRRB_plots_lm, fig.height = 5.5, fig.width = 6, eval=FALSE}


# plots with no facet labels
ggplot(data = plotDataRRBLong,
      aes(x = year, y = value, colour=variable)) +
  geom_line() + 
  # add linear model
  #stat_smooth_func(geom="text",method="lm",hjust=0,parse=TRUE, 
  #                xpos = 2004, ypos = 5800) +
  # add trendline
  stat_smooth(method = 'lm', se=F) +
  scale_colour_manual(values=c('black','red','blue')) + 
  facet_wrap(~variable, nrow=3, scales='free_y') +
  scale_x_continuous(name = '', 
                     breaks = c(2000,2002,2004,2006,2008,2010,2012), 
                     labels = c(2000,2002,2004,2006,2008,2010,2012)) +
  ylab('') +
  theme_bw() + theme(legend.position = 'none',
                     strip.background=element_blank(),
                     strip.text.x = element_blank(),
                     text = element_text(size = 16))


```

## Scatterplots

### Area vs Precip

```{r precipVarea, fig.width = 3, fig.height = 3.25}
# ggplot(data = plotDataRRB,
#        aes(x = zPrecipitation, y = Irrigated_Area)) +
#   geom_point() +
#   ylab('Irrigated Area (km^2)') + xlab('Precipitation (mm)') +
#   ggtitle('Precip vs Irrigated Area') +
#   theme_bw()
```

### Pumping vs precip

```{r precipVpumping, fig.width = 3, fig.height = 3.25}
# ggplot(data = plotDataRRB,
#        aes(x = zPrecipitation, y = Pumping)) +
#   geom_point() +
#   ylab('Pumping (km^3)') + xlab('Precipitation (mm)') +
#   ggtitle('Precip vs Pumping') +
#   theme_bw()
```

### Pumping vs irrigated

```{r areaVpumping, fig.width = 3, fig.height = 3.25}
# ggplot(data = plotDataRRB,
#        aes(x = Irrigated_Area, y = Pumping)) +
#   geom_point() +
#   ylab('Pumping (km^3)') + xlab('Irrigated Area (km^2)') +
#   ggtitle('Irrigated Area vs Pumping') +
#   theme_bw()
```

