---
title: "Irrigation Over Time"
author: "Jill Deines"
date: "December 6, 2016"
output: 
  html_document:
    toc: yes
---

Goal: Investigate trends in irrigated area over time - total area, areas gained, areas lost.

```{r knitrOpts, echo=FALSE}
library(knitr)
opts_chunk$set(cache=FALSE, 
               fig.path='figure/6.00_irrTrends_test4annual/')
```

**R Packages Needed**

```{r packages, message=FALSE, echo=TRUE}
library(rgdal)
library(latticeExtra)
library(tidyverse) # for ggplot2, tidyr, dplyr
library(broom)
library(rgdal)
library(colorRamps)
source('functions/stat_smooth_func_all.R')
```

## Load Data

### RS estimates of irrigated area
Counts of each class type were tallied for several administrative boundaries in GEE

```{r loadGee}
cellArea <- 30*30 # m^2

regionsDir <- 'C:/Users/deinesji/Google Drive/GEE_tableExports/RRB_test4_regional_county_Stats'

# get filenames
files <- list.files(regionsDir, pattern="*_test4_randFor_interannual_plusAncillary.csv")

# read in to a single dataframe
all1 = as.data.frame(do.call(rbind, lapply(files, function(x) {
              csv1 <- read.csv(paste0(regionsDir,'/',x))
              csv1$year <- as.numeric(substr(x, start=1,stop=4))
              return(csv1)
            })))

# convert cell count to area
all1$irrigated_m2 <- all1$X1 * cellArea
all1$irrigated_km2 <- all1$irrigated_m2 / 1000000

# find total #cells in the rrcaExtent
#all1$totalCells <- all1$X0 + all1$X1 + all1$null
#all1[all1$masterID == 'rrcaRrbBuffExtent','totalCells'][1]
```

### County Trends
run regressions to get change in area over time (slope) per county

```{r countyRegress, fig.width = 3.5, fig.height = 3}
# extract county data
counties0 <- all1[grepl('county', all1$masterID),]
counties0$fips5 <- substr(counties0$masterID, start = 11, stop = 15)

# drop columns
counties <- counties0[,c('masterID','fips5','year','irrigated_m2','irrigated_km2')]

# Co county 08041 has NA's; remove
counties <- counties[complete.cases(counties),]

# run regressions by county
countylms <- counties %>% 
  group_by(fips5) %>% 
  do(model = lm(irrigated_km2 ~ year, data = .))

# tidy up with broom
coeffs <- countylms %>% tidy(model)
coeffs <- coeffs[coeffs$term == 'year',]

# map the year coefficient
# get county bounds
gisDir <- "C:/Users/deinesji/Dropbox/1PhdJill/hpa/irrigation/data/GIS/kml/shapefile"
cnty <- readOGR(gisDir, 'Counties_RRCAextent_Clip', verbose=F)

# get state bounds
gisDir2 <- "C:/Users/deinesji/Dropbox/1PhdJill/hpa/irrigation/data/GIS/kml"
states <- readOGR(paste0(gisDir2,'/RRB_fullExtent_states.kml'), verbose=F)
states <- spTransform(states, CRS(proj4string(cnty)))

# join the regression slopes
countymap <- merge(cnty, coeffs[,c('fips5','estimate','p.value')])

# or just significant trends
countymap[is.na(countymap$p.value),'p.value'] <- 1 # dummy p value no data county
countymap$sigCoeffs <- countymap$estimate
countymap[countymap$p.value >= 0.05, 'sigCoeffs'] <- NA

# plot!
pal <- matlab.like(16)
spplot(countymap, 'sigCoeffs', col.regions=pal) +
  latticeExtra::layer(sp.polygons(states, col='black', lwd=2))

# give a starkly different color for declining counties
origPal <- pal
origPal[1] <- 'black'
spplot(countymap, 'sigCoeffs', col.regions=origPal, colorkey = list(space='bottom')) #+
 # latticeExtra::layer(sp.polygons(states, col='black', lwd=2))

# a map with all counties regardless of significance
origPal2 <- origPal
origPal2[2] <- 'black'
spplot(countymap, 'estimate', col.regions=origPal2) +
  latticeExtra::layer(sp.polygons(states, col='black', lwd=2))
```

Which counties have greatest variance over time?

```{r countyVar}
# which counties are most stable?





```


### Novel/Retired Irrigated Area by year

```{r loadYearly}
cellArea <- 30*30 # m^2

yearsDir <- 'C:/Users/deinesji/Google Drive/GEE_tableExports/earliestLatestTable'

# load 
early <- read.csv(paste0(yearsDir, '/', 'test4_randFor_earliest_regionCounts.csv'))
late <- read.csv(paste0(yearsDir, '/', 'test4_randFor_latest_regionCounts.csv'))

# remove county data
early <- early[!grepl('county', early$masterID),]
late <- late[!grepl('county', late$masterID),]

# gather year-county columns into 1 column
early.long <- gather(early, key = year, value = count, X1999:X2015)
late.long <- gather(late, key = year, value = count, X2000:X2016)

# remove junk columns
early.long <- early.long[,c('masterID','year','count')]
late.long <- late.long[,c('masterID','year','count')]

# convert year column to numeric
early.long$year <- as.numeric(substr(early.long$year, 2,5))
late.long$year <- as.numeric(substr(late.long$year, 2,5))

# convert cell count to area
early.long$area_m2 <- early.long$count * cellArea
early.long$area_km2 <- early.long$area_m2 / 1000000

late.long$area_m2 <- late.long$count * cellArea
late.long$area_km2 <- late.long$area_m2 / 1000000
```

### Combine datasets
Make a long format data frame with total irrigated area, novel area each year, lost area each year

```{r combine}
# reformat total area data ------------------
all <- all1[,c('masterID','year','irrigated_m2','irrigated_km2')]
names(all)[3:4] <- c('area_m2', 'area_km2')
all$dataset <- 'total'

# reformat novel/retired data ----------------
# add dataset ID column
early.long$dataset <- 'novel'
late.long$dataset <- 'retired'

columnsWanted <- c('masterID', 'year', 'area_m2', 'area_km2', 'dataset')
# early, drop first years
early2 <- early.long[!(early.long$year %in% c(1999,2000,2001)),
                       columnsWanted]

# late: drop last years
late2 <- late.long[!(late.long$year %in% c(2014,2015,2016)),
                       columnsWanted]

combined <- rbind(all, early2, late2)
```



## Plot Novel/Retired over time

### full study extent
RRB + RRCA + buffer

```{r fullExtent_novelRetired}
ggplot(data = combined[combined$masterID == 'rrcaRrbBuffExtent',],
       aes(x = year, y = area_km2, colour = dataset)) +
    geom_line() + geom_point() +
    ylab(expression(paste("Irrigated Area (", km^2," )"))) +
    xlab('Year') + ggtitle('Full Region') + 
    theme_bw()

ggplot(data = combined[grepl('full', combined$masterID),],
       aes(x = year, y = area_km2, colour = dataset)) + 
    facet_wrap( ~ masterID) +
    geom_line() + geom_point() +
    ylab(expression(paste("Irrigated Area (", km^2," )"))) +
    xlab('Year') + ggtitle('Full Region by State') +
    theme_bw()
```

Dave and Anthony requested a plot with cumulative novel and retired areas

```{r fullExtent_novelRetired_cumulative, fig.width = 6.5}
# extract dataset: full study extent, by state
stateExtent <- combined[grepl('full', combined$masterID),]



# make a cumulative sum column by dataset type
stateExtent <- stateExtent %>%
  group_by(masterID, dataset) %>%
  arrange(year) %>%
  mutate(cumArea_km2 = cumsum(area_km2))

# make a column to plot that has raw data for total, cumulative for novel/retired
stateExtent$toPlot <- NA
stateExtent[stateExtent$dataset == 'total','toPlot'] <- stateExtent[stateExtent$dataset == 'total','area_km2']
stateExtent[stateExtent$dataset == 'novel','toPlot'] <- stateExtent[stateExtent$dataset == 'novel','cumArea_km2']
stateExtent[stateExtent$dataset == 'retired','toPlot'] <- stateExtent[stateExtent$dataset == 'retired','cumArea_km2']

# plotty plot
labels <- c(`CO_full_RRCA` = 'Colorado',
            `KS_full_RRCA` = 'Kansas',
            `NE_full_RRCA` = 'Nebraska')


ggplot(data = stateExtent,
       aes(x = year, y = toPlot, colour = dataset)) + 
    facet_wrap( ~ masterID, labeller = as_labeller(labels)) +
    geom_line() + geom_point() +
    ylab(expression(paste("Irrigated Area (", km^2,")"))) +
    xlab('Year') + #ggtitle('Full Region by State') +
    scale_color_discrete(breaks=c('total','novel','retired'),
                         labels=c('total', 'cumulative novel',
                                  'cumulative retired')) +
    guides(color=guide_legend(title=NULL)) +
    theme_bw() +
    theme(legend.position = 'top') 

ggplot(data = stateExtent,
       aes(x = year, y = toPlot, colour = dataset)) + 
    facet_wrap( ~ masterID, labeller = as_labeller(labels)) +
    geom_line() + geom_point() +
    ylab(expression(paste("Irrigated Area (", km^2,")"))) +
    xlab('Year') + #ggtitle('Full Region by State') +
    scale_color_discrete(breaks=c('total','novel','retired'),
                         labels=c('total', 'cumulative novel',
                                  'cumulative retired')) +
    guides(color=guide_legend(title=NULL)) +
    theme_bw() +
    theme(legend.position = c(.85,.5)) 
```

And a plot of just retired/novel fields over time
```{r fullExtent_novelRetired_alone, fig.height = 2.5, fig.width = 6.5}
# remove total from dataset
stateExtent2 <- stateExtent[stateExtent$dataset != 'total',]

# plotty plot
labels <- c(`CO_full_RRCA` = 'Colorado',
            `KS_full_RRCA` = 'Kansas',
            `NE_full_RRCA` = 'Nebraska')

ggplot(data = stateExtent2,
       aes(x = year, y = area_km2, colour = dataset)) + 
    facet_wrap( ~ masterID, labeller = as_labeller(labels)) +
    geom_line() + geom_point() +
    ylab(expression(paste("Irrigated Area (", km^2,")"))) +
    xlab('Year') + #ggtitle('Full Region by State') +
    guides(color=guide_legend(title=NULL)) +
    theme_bw() +
    theme(legend.position = c(0.15,.6)) 
```


### basin only
rrb basin bound

```{r basin_novelRetired}


ggplot(data = combined[grepl('basin', combined$masterID) &
                     !grepl('NRD', combined$masterID),],
       aes(x = year, y = area_km2, colour = dataset)) + 
    facet_wrap( ~ masterID) +
    geom_line() + geom_point() +
    ylab(expression(paste("Irrigated Area (", km^2," )"))) +
    xlab('Year') + ggtitle('RRB Basin by State') +
    theme_bw()
```


basin level cumulative

```{r basinExtent_novelRetired_cumulative, fig.width = 6.5}
# extract dataset: full study extent, by state
stateBasin <- combined[grepl('basin', combined$masterID) &
                     !grepl('NRD', combined$masterID),]

# make a cumulative sum column by dataset type
stateBasin <- stateBasin %>%
  group_by(masterID, dataset) %>%
  arrange(year) %>%
  mutate(cumArea_km2 = cumsum(area_km2))

# make a column to plot that has raw data for total, cumulative for novel/retired
stateBasin$toPlot <- NA
stateBasin[stateBasin$dataset == 'total','toPlot'] <- stateBasin[stateBasin$dataset == 'total','area_km2']
stateBasin[stateBasin$dataset == 'novel','toPlot'] <- stateBasin[stateBasin$dataset == 'novel','cumArea_km2']
stateBasin[stateBasin$dataset == 'retired','toPlot'] <- stateBasin[stateBasin$dataset == 'retired','cumArea_km2']

labels <- c(`CO_basin` = 'Colorado',
            `KS_basin` = 'Kansas',
            `NE_basin` = 'Nebraska')

# plotty plot
ggplot(data = stateBasin,
       aes(x = year, y = toPlot, colour = dataset)) + 
    facet_wrap( ~ masterID, labeller = as_labeller(labels)) +
    geom_line() + geom_point() +
    ylab(expression(paste("Irrigated Area (", km^2,")"))) +
    xlab('Year') + #ggtitle('Full Region by State') +
    scale_color_discrete(breaks=c('total','novel','retired'),
                         labels=c('total', 'cumulative novel',
                                  'cumulative retired')) +
    guides(color=guide_legend(title=NULL)) +
    theme_bw() +
    theme(legend.position = 'top') 
```

## Plot Area Data (plots prior to novel/retired analysis)

### Full RRCA RRB Extent

```{r rrcaExtentPlots, fig.height=2.5}
ggplot(data = all1[all1$masterID == 'rrcaRrbBuffExtent',],
       aes(x = year, y = irrigated_km2)) +
    geom_line() + geom_point() +
    ylab(expression(paste("Irrigated Area (", km^2," )"))) +
    xlab('Year') + ggtitle('Full Region') + 
    theme_bw()
# precip
ggplot(data = all1[all1$masterID == 'rrcaRrbBuffExtent',],
       aes(x = year, y=pr_ann)) +
    geom_line() + geom_point() +
    ylab("Dec - Aug precip  (mm)") +
    xlab('Year') + ggtitle('Full Region') + 
    theme_bw()

# aridity
ggplot(data = all1[all1$masterID == 'rrcaRrbBuffExtent',],
       aes(x = year, y=aridity)) +
    geom_line() + geom_point() +
    ylab("Aridity") +
    xlab('Year') + ggtitle('Full Region') + 
    theme_bw()
```

```{r rrcabyState}

ggplot(data = all1[grepl('full', all1$masterID),],
       aes(x = year, y = irrigated_km2, colour = masterID)) +
    geom_line() + geom_point() +
    ylab(expression(paste("Irrigated Area (", km^2," )"))) +
    xlab('Year') + ggtitle('Full Region by State') +
    theme_bw()
```

```{r rrcaStateTrends}
stateRRCA <- all1[grepl('full', all1$masterID),]
stateRRCA$masterID <- as.character(stateRRCA$masterID)

# run regressions by state 
statelms <- stateRRCA %>% 
  group_by(masterID) %>% 
  do(model = lm(irrigated_km2 ~ year, data = .))

# tidy up with broom
coeffs2 <- statelms %>% tidy(model)
coeffs2 <- coeffs2[coeffs2$term == 'year',]

# plots with no facet labels
ggplot(data = stateRRCA,
      aes(x = year, y = irrigated_km2, colour=masterID)) +
  geom_line() + 
  # add linear model
  #stat_smooth_func(geom="text",method="lm",hjust=0,parse=TRUE, 
  #                xpos = 2000, ypos = 14000) +
  # add trendline
  stat_smooth(method = 'lm', se=T) +
  ylab('') +
  theme_bw() + theme(legend.position = 'none',
                     strip.background=element_blank(),
                     strip.text.x = element_blank(),
                     text = element_text(size = 16))

```


### RRB Basin plots

```{r rrbBasinPlots}
ggplot(data = all1[grepl('basin', all1$masterID) &
                     !grepl('NRD', all1$masterID),],
       aes(x = year, y = irrigated_km2, colour = masterID)) +
    geom_line() + geom_point() +
    ylab(expression(paste("Irrigated Area (", km^2," )"))) +
    xlab('Year') + ggtitle('RRB by State') +
   theme_bw() 
```

```{r neBasin, fig.height=2.5}
ggplot(data = all1[grepl('NE_basin', all1$masterID),],
       aes(x = year, y = irrigated_km2)) +
    geom_line() + geom_point() +
    ylab(expression(paste("Irrigated Area (", km^2," )"))) +
    xlab('Year') + ggtitle('Nebraska RRB') +
   theme_bw() 
```
### NRD Basin plots

```{r nrdBasinPlots}
ggplot(data = all1[grepl('NRD', all1$masterID) | 
                     grepl('GMD', all1$masterID),],
       aes(x = year, y = irrigated_km2, colour = masterID)) +
    geom_line() + geom_point() +
    ylab(expression(paste("Irrigated Area (", km^2," )"))) +
    xlab('Year') + ggtitle('By NE NRD') +
   theme_bw() 

```



