---
title: "Statistical Model"
author: "Jill Deines"
date: "February 27, 2017"
output: 
  html_document:
    toc: yes
---

Goal: Investigate relationship among irrigated area, precip, pumping volume, and crop prices over time 

```{r knitrOpts, echo=FALSE}
library(knitr)
opts_chunk$set(cache=FALSE, 
               fig.path='figure/6.05_statModels/')
```

**R Packages Needed**

```{r packages, message=FALSE, echo=TRUE}
library(ggplot2)
```

## Data wrangling

### NASS crop price data

First, make a table of inflation adjustments relative to 2015, the final year available in my NASS price dataset. Write out so this is static and not dependent on external download

```{r inflationTable, eval=FALSE}
# adjust for inflation
library(lubridate) 
library(dplyr) 

# download the consumer price index and get adjustments 
monthly_cpi <- read.csv("http://research.stlouisfed.org/fred2/data/CPIAUCSL.csv",
                        header = TRUE)
monthly_cpi$cpi_year <- year(monthly_cpi$DATE)
yearly_cpi <- monthly_cpi %>% group_by(cpi_year) %>% summarize(cpi = mean(VALUE))
yearly_cpi$adj_factor_2015 <- yearly_cpi$cpi/yearly_cpi$cpi[yearly_cpi$cpi_year == 2015]
# for posterity
yearly_cpi$adj_factor_2016 <- yearly_cpi$cpi/yearly_cpi$cpi[yearly_cpi$cpi_year == 2016]
yearly_cpi$adj_factor_2017 <- yearly_cpi$cpi/yearly_cpi$cpi[yearly_cpi$cpi_year == 2017]

# write out
nassdir <- 'C:/Users/deinesji/Dropbox/1PhdJill/hpa/irrigation/data/nass/price'
write.csv(yearly_cpi, paste0(nassdir, '/cpi_priceAdjustments.csv'), 
          row.names=F)
```

Price data was downloaded manually from the NASS quickstats tool (https://quickstats.nass.usda.gov/) for the following query: SURVEY - CROPS - FIELD CROPS - CORN/SOYBEANS/WHEAT/HAY/SORGHUM - PRICE RECEIVED - 

* CORN, GRAIN - PRICE RECEIVED, MEASURED IN $ / BU
* HAY, PRICE RECEIVED, MEASURED IN $ / TON (short ton)
* SORGHUM, GRAIN - PRICE RECEIVED, MEASURED IN $ / CWT (hundredweight)
* SOYBEANS - PRICE RECEIVED, MEASURED IN $ / BU
* WHEAT - PRICE RECEIVED, MEASUTED IN $ / BU

- TOTAL - NATIONAL - US TOTAL - 1996:2015 - ANNUAL - YEAR


Here, I load the data and clean it to have consistent pricing across crops and years

```{r nassPrice}
# load nass
nassdir <- 'C:/Users/deinesji/Dropbox/1PhdJill/hpa/irrigation/data/nass/price'
price0 <- read.csv(paste0(nassdir, '/nass_prices_5crops_national_1996-2015.csv'))

# load consumer price index adjustments
cpi <- read.csv(paste0(nassdir, '/cpi_priceAdjustments.csv'))

# make a units columns
unitkey <- data.frame(Commodity = c('CORN','HAY','SORGHUM','SOYBEANS','WHEAT'),
                      unit = c('BU','TON','CWT','BU','BU'))
price0 <- merge(price0,unitkey)

# remove silly columns
price0 <- price0[,c('Commodity','Year','Value','unit')]

# add the adjustment factor
price <- merge(price0, cpi[,c('cpi_year','adj_factor_2015')], by.x = 'Year', 
               by.y = 'cpi_year')

# adjust prices
price$price_2015 <- price$Value / price$adj_factor_2015

# plot it - unadjusted
ggplot(data = price, aes(x=Year, y = Value)) +
  geom_line() +
  facet_wrap(~Commodity, scales='free_y') + theme_bw() +
  ggtitle('Nomial (unadjusted) Crop Prices') 


# plot it
ggplot(data = price, aes(x=Year, y = price_2015)) +
  geom_line() +
  facet_wrap(~Commodity, scales='free_y') + theme_bw() +
  ggtitle('Inflation-adjusted Crop Prices')

# extract only corn
cornprice <- price[price$Commodity == 'CORN',]

ggplot(data = cornprice, aes(x=Year, y = price_2015)) +
  geom_line() + theme_bw() +
  ggtitle('Inflation-adjusted Corn Prices')
```

