---
title: "Statistical Model"
author: "Jill Deines"
date: "July 6, 2017"
output: 
  html_document:
    toc: yes
---

Goal: Investigate relationship among irrigated area, precip, pumping volume, and crop prices over time 

updated from 6.05 Statistical Models, test5


```{r knitrOpts, echo=FALSE}
library(knitr)
opts_chunk$set(cache=FALSE, 
               fig.path='figure/6.06_statModels_test5_rrb/')
```

**R Packages Needed**

```{r packages, message=FALSE, echo=TRUE}
library(gridExtra)
library(grid)
library(tidyverse) # for ggplot2, tidyr, dplyr
library(broom)
library(GGally)
library(corrplot)
library(Hmisc)
library(party) #carts
source('functions/stat_smooth_func_all.R')
library(RColorBrewer)
library(rgdal)
```

## User Vars
Derived dataset source files

* corn prices: produced in first chunks of this script
* RRCA pumping volume: processed in 4.21_Figures_volumeVsArea_randFor... 'loadPumpingData' chunk
* RRB pumping volume from the RRCA; also processed in 4.21


```{r userVars}
# data directory
derivedDir <- 'C:/Users/deinesji/Dropbox/1PhdJill/hpa/irrigation/data/derivedDatasets'
# corn prices
cornFile <- 'price_nass_corn_1996-2016.csv'
# RRCA volume
RRCA.volume.file <- 'RRCA_fullModel_irrigation_volume.csv'
# RRB volume
RRB.volume.file <- 'RRB_byState_irrigation_volume.csv'
```

## Data wrangling
(jump to Analyses/loaddata chunk after first time)

### NASS crop price data

First, make a table of inflation adjustments relative to 2015, the final year available in my NASS price dataset. Write out so this is static and not dependent on external download

```{r inflationTable, eval=FALSE}
# adjust for inflation
library(lubridate) 
library(dplyr) 

# download the consumer price index and get adjustments 
monthly_cpi <- read.csv("http://research.stlouisfed.org/fred2/data/CPIAUCSL.csv",
                        header = TRUE)
monthly_cpi$cpi_year <- year(monthly_cpi$DATE)
yearly_cpi <- monthly_cpi %>% group_by(cpi_year) %>% summarize(cpi = mean(VALUE))
yearly_cpi$adj_factor_2015 <- yearly_cpi$cpi/yearly_cpi$cpi[yearly_cpi$cpi_year == 2015]
# for posterity
yearly_cpi$adj_factor_2016 <- yearly_cpi$cpi/yearly_cpi$cpi[yearly_cpi$cpi_year == 2016]
yearly_cpi$adj_factor_2017 <- yearly_cpi$cpi/yearly_cpi$cpi[yearly_cpi$cpi_year == 2017]

# write out
nassdir <- 'C:/Users/deinesji/Dropbox/1PhdJill/hpa/irrigation/data/nass/price'
write.csv(yearly_cpi, paste0(nassdir, '/cpi_priceAdjustments.csv'), 
          row.names=F)
```

Price data was downloaded manually from the NASS quickstats tool (https://quickstats.nass.usda.gov/) for the following query: SURVEY - CROPS - FIELD CROPS - CORN/SOYBEANS/WHEAT/HAY/SORGHUM - PRICE RECEIVED - 

* CORN, GRAIN - PRICE RECEIVED, MEASURED IN $ / BU
* HAY, PRICE RECEIVED, MEASURED IN $ / TON (short ton)
* SORGHUM, GRAIN - PRICE RECEIVED, MEASURED IN $ / CWT (hundredweight)
* SOYBEANS - PRICE RECEIVED, MEASURED IN $ / BU
* WHEAT - PRICE RECEIVED, MEASUTED IN $ / BU

- TOTAL - NATIONAL - US TOTAL - 1996:2016 - ANNUAL - YEAR


Here, I load the data and clean it to have consistent pricing across crops and years

```{r nassPrice, eval=FALSE}
# load nass
nassdir <- 'C:/Users/deinesji/Dropbox/1PhdJill/hpa/irrigation/data/nass/price'
price0 <- read.csv(paste0(nassdir, '/nass_prices_5crops_national_1996-2016.csv'))

# load consumer price index adjustments
cpi <- read.csv(paste0(nassdir, '/cpi_priceAdjustments.csv'))

# make a units columns
unitkey <- data.frame(Commodity = c('CORN','HAY','SORGHUM','SOYBEANS','WHEAT'),
                      unit = c('BU','TON','CWT','BU','BU'))
price0 <- merge(price0,unitkey)

# remove silly columns
price0 <- price0[,c('Commodity','Year','Value','unit')]

# add the adjustment factor
price <- merge(price0, cpi[,c('cpi_year','adj_factor_2015','adj_factor_2016')], 
               by.x = 'Year', by.y = 'cpi_year')

# adjust prices (note: matched trials at https://data.bls.gov/cgi-bin/cpicalc.pl)
price$price_2015 <- price$Value / price$adj_factor_2015
price$price_2016 <- price$Value / price$adj_factor_2016

# plot it - unadjusted
ggplot(data = price, aes(x=Year, y = Value)) +
  geom_line() +
  facet_wrap(~Commodity, scales='free_y') + theme_bw() +
  ggtitle('Nomial (unadjusted) Crop Prices') 


# plot it
ggplot(data = price, aes(x=Year, y = price_2016)) +
  geom_line() +
  facet_wrap(~Commodity, scales='free_y') + theme_bw() +
  ggtitle('Inflation-adjusted Crop Prices')

# extract only corn
cornprice <- price[price$Commodity == 'CORN',]

ggplot(data = cornprice, aes(x=Year, y = price_2016)) +
  geom_line() + theme_bw() +
  ggtitle('Inflation-adjusted Corn Prices')

# write out dataset
write.csv(cornprice, paste0(derivedDir, '/price_nass_corn_1996-2016.csv'), 
          row.names=F)
```

## Analyses

### Load Data
loads datasets produced/cleaned from other scripts

```{r loadData}
# usda nass annual corn price, $ / bushel
cornprice <- read.csv(paste0(derivedDir, '/',cornFile))

# irrigation volume ----------
# full rrca
volume.state <- read.csv(paste0(derivedDir, '/',RRCA.volume.file))
# aggregate volume across states - sum
volume <- aggregate(total_km3 ~ year, data = volume.state, FUN = 'sum', na.rm=T)

# rrb portion of the rrca
volume.state.rrb <- read.csv(paste0(derivedDir, '/',RRB.volume.file))
# aggregate volume across states - sum
volume.rrb <- aggregate(total_km3 ~ year, data = volume.state, FUN = 'sum', na.rm=T)

```

### load Irrigated Area and Precip
Extract irrigated area and precip for regions of interests

* RRCA domain irrigation and precip in 4.21_Figures_volumeVsArea_randFor... in the 'loadAreaData' chunk

```{r irrData}
# load REGIONAL AREA + precip data files ----------------------------------------
cellArea <- 30*30 # m^2
regionsDir <- 'C:/Users/deinesji/Google Drive/GEE_tableExports/RRB_test5_regional_county_Stats'

# get filenames
files <- list.files(regionsDir,
    pattern="*_test5_randFor_interannual_plusAncillary_REVISION_BOUNDARIES.csv")

# read in to a single dataframe (function)
loadStats <- function(fileList){
  datOut = as.data.frame(do.call(rbind, lapply(fileList, function(x) {
              csv1 <- read.csv(paste0(regionsDir,'/',x))
              csv1$year <- as.numeric(substr(x, start=1,stop=4))
              return(csv1)
           })))  
  # convert cell count to area
  datOut$irrigated_m2 <- datOut$X1 * cellArea
  datOut$irrigated_km2 <- datOut$irrigated_m2 / 1000000
  return(datOut)
}

# load
all1 = loadStats(files)
# remove county data 
all1 <- all1[!grepl('county', all1$masterid),]
# drop a bunch of columns
all <- all1[,c('masterid','year','irrigated_km2','pr_ann')]

# extract regions of interest
regionsWanted_rrbrrca <- c('RRCARRB_KS','RRCARRB_CO','RRCARRB_NE')
areappt.rrbstates <- all[all$masterid %in% regionsWanted_rrbrrca,]

# extract regions of interest
areappt.rrb <- all[all$masterid == 'RRCARRB_intersection',]
```

### Merge datasets together

```{r mergeData}
# merge price
areappt.price.rrb <- merge(areappt.rrb, cornprice[,c('Year','price_2016')], 
                        by.x = 'year', by.y = 'Year')
areappt.price.rrbstates <- merge(areappt.rrbstates, cornprice[,c('Year','price_2016')], 
                        by.x = 'year', by.y = 'Year')

# merge volume
rrca.rrb <- merge(areappt.price.rrb,volume.rrb, all=T)

# make state key
areappt.price.rrbstates$key <- paste0(substr(areappt.price.rrbstates$masterid,9,10),
                                      areappt.price.rrbstates$year)
rrca.rrb.states <- merge(areappt.price.rrbstates, volume.state.rrb, all.x=T, by='key')

# rename columns
names(rrca.rrb) <- c('Year','masterid','Area_km2','Precip_mm','Price','Volume_km3')
names(rrca.rrb.states)[c(2:6,9)] <- c('Year','masterid','Area_km2','Precip_mm',
                                    'Price','Volume_km3') 
rrca.rrb.states <- rrca.rrb.states[,c('Year','masterid','Area_km2','Precip_mm',
                                    'Price','Volume_km3')]

# add a price lag
rrca.rrb$priceLag1 <- cornprice$price_2016[3:20]

cornprice$priceLag1 <- c(NA,cornprice$price_2016[1:20]) # so merge has correct column name
rrca.rrb.states <- merge(rrca.rrb.states,cornprice[,c('Year', 'priceLag1')],all.x=T)

# add total irrigated area for the full study region
rrca.rrb$area.union <- all[all$masterid == 'RRCA_RRB_Union','irrigated_km2']
rrca.rrb$Area_km2_else <- rrca.rrb$area.union - rrca.rrb$Area_km2
```

### Exploratory analyses: RRB of RRCA domain

Average application depth: Divide irrigated pumping by area (average application depth); plot vs precip expressed in mm

what proportion of water demand they are satisfying?

```{r applicationDepth_cm_noStats_rrbrrca, fig.height = 3, fig.width = 3, dpi=600, dev=c('png','pdf')}
# calculate depth of water applied
rrca.rrb$Depth_m <- rrca.rrb$Volume_km3/rrca.rrb$Area_km2 * 1000

# convert dpeth and precip to cm
rrca.rrb$Depth_cm <- rrca.rrb$Depth_m * 100
rrca.rrb$Precip_cm <- rrca.rrb$Precip_mm / 10

depth <- ggplot(rrca.rrb, aes(Precip_cm, Depth_cm)) +
  geom_point(color = 'chartreuse3') + theme_bw() +
  stat_smooth(method = 'lm', se=F, size=.5, color='chartreuse3') + 
  #stat_smooth_func_all(geom="text",method="lm",hjust=0,parse=TRUE, 
  #               xpos = 25, ypos = 68) +
  ylab('Depth (cm)') + xlab('Precipitation (cm)')  +
  theme(axis.text=element_text(size=10),
               legend.text=element_text(size=10),
               axis.title=element_text(size=11),        
            panel.grid.major = element_blank(),
             panel.grid.minor = element_blank())
depth

depthlm <- lm(Depth_cm ~ Precip_cm, data = rrca.rrb)
summary(depthlm)

```

by state?

```{r applicationDepth_cm_noStats_rrbrrcaState, fig.height = 3, fig.width = 7, dpi=600, dev=c('png','pdf')}
# calculate depth of water applied
rrca.rrb.states$Depth_m <- rrca.rrb.states$Volume_km3/rrca.rrb.states$Area_km2 * 1000

# convert dpeth and precip to cm
rrca.rrb.states$Depth_cm <- rrca.rrb.states$Depth_m * 100
rrca.rrb.states$Precip_cm <- rrca.rrb.states$Precip_mm / 10

depth.state <- ggplot(rrca.rrb.states, aes(Precip_cm, Depth_cm)) +
  geom_point(color = 'chartreuse3') + theme_bw() +
  facet_wrap(~masterid)+
  stat_smooth(method = 'lm', se=F, size=.5, color='chartreuse3') + 
  stat_smooth_func_all(geom="text",method="lm",hjust=0,parse=TRUE, 
                 xpos = 20, ypos = 20) +
  ylab('Depth (cm)') + xlab('Precipitation (cm)')  +
  theme(axis.text=element_text(size=10),
               legend.text=element_text(size=10),
               axis.title=element_text(size=11),        
            panel.grid.major = element_blank(),
             panel.grid.minor = element_blank())
depth.state
```

And plot all time series.

```{r timeSeriesPlot_finalish_rrbrrca, fig.height = 4, fig.width = 6, dpi=600, dev=c('png','pdf')}
# data frame for plotting
rrcaPlot <- rrca.rrb

# items to plot
names(rrcaPlot)[6] <- 'dVolume_km3'
plotUs <- c('Area_km2','dVolume_km3','Precip_cm','Price')
#plotUs <- c('AArea_km2_else','Volume_km3','Precip_cm','Price')

# gather
plotdata <- gather_(rrcaPlot, key_col = 'variable', value_col = 'value',
                          gather_cols = plotUs)

# plots with trendlines no text
ggplot(data = plotdata,
      aes(x = Year, y = value, colour=variable)) +
  geom_line() + geom_point() +
  geom_line(data = data.frame(rrca.rrb,variable='Area_km2'), 
            aes(x = Year, y = Area_km2_else), color='gray55') +
  geom_point(data = data.frame(rrca.rrb,variable='Area_km2'), 
            aes(x = Year, y = Area_km2_else),color='gray55') +
  stat_smooth(data = data.frame(rrca.rrb,variable='Area_km2'), 
            aes(x = Year, y = Area_km2_else), method='lm', 
            se=F,size=.5, color='gray55') +
  scale_colour_manual(values=c('black','red','blue','orange')) + 
  facet_wrap(~variable, nrow=5, scales='free_y') +
  stat_smooth(data = subset(plotdata, variable %in% 
                              c('Area_km2','Price')), 
              method = 'lm', se=F, size=.5) +
  scale_x_continuous(name = '', 
                  breaks = c(2000,2002,2004,2006,2008,2010,2012,2014,2016), 
                 labels = c(2000,2002,2004,2006,2008,2010,2012,2014,2016)) +
  ylab('') +
  theme_bw() + theme(legend.position = 'none',
                     strip.background=element_blank(), # comment to add labels
                     strip.text.x = element_blank(),   # comment to add labels
                     text = element_text(size = 15),
                                          panel.grid.major = element_blank(),
             panel.grid.minor = element_blank())

```

Run a quick set of individual regressions on these time series

```{r timeSeriesLMs}
# run regressions by county
lms <- plotdata %>% 
  group_by(variable) %>% 
  do(model = lm(value ~ Year, data = .))

# tidy up with broom: model coefficients
coeffs <- lms %>% tidy(model)
coeffs <- coeffs[coeffs$term == 'Year',]
coeffs

# and model R2 values
rsq <- lms %>% glance(model)
rsq

# Unioned area outside of the RRB/RRCA intersection
summary(lm(Area_km2_else ~ Year,data=rrca.rrb))
```

#### redo with percent area
Dave/Anthony suggested using percent of study area irrigated instead of absolute area

Get relevant boundary areas
```{r getAreas}
gisDir <- 'C:/Users/deinesji/Dropbox/1PhdJill/hpa/irrigation/data/GIS/kml/FINAL_BOUNDARIES'

# rrb rrca intersection
poly.rrca.rrb.state <- readOGR(gisDir, 'RRCA-RRB_interesection_byState_aea', verbose = F)
area.rrca.rrb <- sum(poly.rrca.rrb.state$areakm2)

# full rrca rrb union
poly.union <- readOGR(gisDir, 'RRCA_RRB_union_aea', verbose = F)
area.union <- poly.union$areakm2

# non rrca/rrb intersect union
elseArea <- area.union - area.rrca.rrb
```

and plot percentages

```{r timeSeriesPlot_percent_finalish_rrcarrb, fig.height = 4, fig.width = 6, dpi=600, dev=c('png','pdf')}
# get percent of RRB*RRCA area irrigated
rrca.rrb$AreaPercentIn <- (rrca.rrb$Area_km2/area.rrca.rrb) * 100

# get percetn of non RRB*RRCA area irrigated
rrca.rrb$AreaPercentOut <- (rrca.rrb$Area_km2_else/elseArea) * 100

# items to plot
names(rrca.rrb)[6] <- 'dVolume_km3'

# items to plot
plotUs2 <- c('AreaPercentIn','dVolume_km3','Precip_cm','Price')

# gather
plotdata2 <- gather_(rrca.rrb, key_col = 'variable', value_col = 'value',
                          gather_cols = plotUs2)


# plots with trendlines no text, no outside region line
ptime <- ggplot(data = plotdata2,
      aes(x = Year, y = value, colour=variable)) +
  geom_line() + geom_point() +
    scale_colour_manual(values=c('black','red','blue','orange')) + 
  facet_wrap(~variable, nrow=4, scales='free_y') +
  stat_smooth(data = subset(plotdata2, variable %in% 
                              c('AreaPercentIn','Price')), 
              method = 'lm', se=F, size=.4) +
  # add second line to panel a
  # geom_line(data = data.frame(rrca.rrb,variable='AreaPercentIn'), 
  #           aes(x = Year, y = AreaPercentOut), color='gray55') +
  # geom_point(data = data.frame(rrca.rrb,variable='AreaPercentIn'), 
  #           aes(x = Year, y = AreaPercentOut),color='gray55') +
  # stat_smooth(data = data.frame(rrca.rrb,variable='AreaPercentIn'), 
  #           aes(x = Year, y = AreaPercentOut), method='lm', 
  #           se=F,size=.5, color='gray55') +
  scale_x_continuous(name = '', 
                  breaks = c(2000,2002,2004,2006,2008,2010,2012,2014,2016), 
                 labels = c(2000,2002,2004,2006,2008,2010,2012,2014,2016)) +
  ylab('') + xlab('Year') +
  theme_bw() + theme(legend.position = 'none',
                     strip.background=element_blank(), # comment to add labels
                     strip.text.x = element_blank(),   # comment to add labels
                      axis.text=element_text(size=10),
                     legend.text=element_text(size=10),
                     axis.title=element_text(size=11),
                             panel.grid.major = element_blank(),
             panel.grid.minor = element_blank())
ptime

# plots with trendlines no text, with outside region line
ptime2 <- ggplot(data = plotdata2,
      aes(x = Year, y = value, colour=variable)) +
 # add second line to panel a
  geom_line(data = data.frame(rrca.rrb,variable='AreaPercentIn'),
            aes(x = Year, y = AreaPercentOut), color='gray55') +
  geom_point(data = data.frame(rrca.rrb,variable='AreaPercentIn'),
            aes(x = Year, y = AreaPercentOut),color='gray55') +
  stat_smooth(data = data.frame(rrca.rrb,variable='AreaPercentIn'),
            aes(x = Year, y = AreaPercentOut), method='lm',
            se=F,size=.5, color='gray55') +
   geom_line() + geom_point() +
    scale_colour_manual(values=c('black','red','blue','orange')) + 
  facet_wrap(~variable, nrow=4, scales='free_y') +
  stat_smooth(data = subset(plotdata2, variable %in% 
                              c('AreaPercentIn','Price')), 
              method = 'lm', se=F, size=.4) +
  
  scale_x_continuous(name = '', 
                  breaks = c(2000,2002,2004,2006,2008,2010,2012,2014,2016), 
                 labels = c(2000,2002,2004,2006,2008,2010,2012,2014,2016)) +
  ylab('') + xlab('Year') +
  theme_bw() + theme(legend.position = 'none',
                     strip.background=element_blank(), # comment to add labels
                     strip.text.x = element_blank(),   # comment to add labels
                      axis.text=element_text(size=10),
                     legend.text=element_text(size=10),
                     axis.title=element_text(size=11),
                             panel.grid.major = element_blank(),
             panel.grid.minor = element_blank())
ptime2
```

Run a quick set of individual regressions on these time series

```{r timeSeriesLMs_percent}
# run regressions by county
lms <- plotdata2 %>% 
  group_by(variable) %>% 
  do(model = lm(value ~ Year, data = .))

# tidy up with broom: model coefficients
coeffs <- lms %>% tidy(model)
coeffs <- coeffs[coeffs$term == 'Year',]
coeffs

# and model R2 values
rsq <- lms %>% glance(model)
rsq

# Unioned area outside of the RRB/RRCA intersection
summary(lm(AreaPercentOut ~ Year,data=rrca.rrb))
```

#### Correlations
Let's look at some preliminary correlations for the RRCA domain. 

```{r corrPlots_rrbrrca, fig.height=5, fig.width=5, dpi=600, dev=c('png','pdf')}
# dataset to use
rrcarrb <- rrca.rrb

# remove the full domain data and other columns not to include in cors
nonsense <- c('masterid', 'Depth_m', 'Precip_mm','area.union', 'Area_km2_else',
              'AreaPercentIn','AreaPercentOut')
rrcarrb2 <- rrcarrb[,-which(names(rrcarrb) %in% nonsense)]

# give legible names
names(rrcarrb2) <- c('Year','Area','Price','Volume','Price_Lag','Depth','Precip')

# order columns alphabetically
rrcarrb2 <- rrcarrb2[,order(names(rrcarrb2))]

# cor plots
varcor <- cor(rrcarrb2, use='complete.obs')
corrplot(varcor, method='color', addCoef.col='black', tl.col='black',
         type = 'lower', 
         col = colorRampPalette(brewer.pal(7,'RdBu'))(200))



# sig levels
corsig <- rcorr(as.matrix(rrcarrb2[complete.cases(rrcarrb2),]), type = 'pearson')
options(scipen = 999)
corsig$P
```

Interpretation:

* Precip has inverse correlations with irrigation volume (significant) and irrigation area (non-significant)
* Volume and area are negatively correlated (p = 0.077)
* Price Lag 1 correlated with area
* Depth negatively corrected with area, precip; positive with volume

```{r scatterplots_rrbrrca, fig.height=5.5, fig.width=8.5}
# an extra function to add trendlines
my_fn <- function(data, mapping, ...){
  p <- ggplot(data = data, mapping = mapping) + 
    geom_point() + 
    geom_smooth(method=loess, fill="red", color="red", ...) +
    geom_smooth(method=lm, fill="blue", color="blue", ...)
  p
}

# scatterplot matrix
ggpairs(rrcarrb2, lower=list(continuous = my_fn)) 
```

Hm. A two step model? Predict application depth from precip; predict irrigated area from depth and price? Or a hierarchical model where depth comes from precip...


### CaRT
Use a CART to analysis, since robust to autocorrelation and we don't have long enough time series for robust time series analysis.

I picked 'ctree' function the party package because of the nice-ish output on an R Blogger's post, but Andy subsequently suggested it as well. Look at that!

From 'party' package

```{r cart_rrbrrca, fig.height = 5, fig.width = 5, dpi=600, dev=c('png','pdf')}
# add percent area back in, to do it that way
rrcarrb2$AreaP <- rrcarrb$AreaPercentIn

# rrcarrb domain dataset - complete across variables
rrcarrb3 <- rrcarrb2[complete.cases(rrcarrb2),]

# add percent area back in, to do it that way

# new cart, RRCA area ---------------------------------------------
 # control tree
controls <- ctree_control(minsplit = 3, minbucket = 3, mincriterion = .85)

# run tree
ct <- ctree(AreaP ~ Precip + Price_Lag, data = rrcarrb3, 
            controls = controls)
plot(ct, main="Conditional Inference Tree")

# predict on same data
predicted <- predict(ct)
ghettoR2 <- cor(as.vector(predicted), rrcarrb3$AreaP)^2
ghettoR2
```

Visualize these splits in space

```{r heatmapCart_rrbrrca, fig.height=3, fig.width = 3.5}
ggplot(rrcarrb3,
       aes(x=Precip, y = Price_Lag, colour = AreaP)) +
  geom_point(size = 3) + 
  geom_hline(yintercept = 3.924, colour='red') + 
  geom_segment(aes(x=46.07, xend = 46.07, y=3.924, yend=max(Price_Lag)), colour = 'black') + 
  ylab('Price Lag ($/bushel)') + xlab('Precip. (cm)') + 
  theme_bw()
         
```

Visualize differently
```{r cartplot2_rrbrrca, fig.height=3, fig.width = 3.5, dpi=600, dev=c('png','pdf')}
# add a factor variable based on price split
rrcarrb3$PriceSplit <- NA
rrcarrb3[rrcarrb3$Price_Lag > 3.92,'PriceSplit'] <- 'High'
rrcarrb3[is.na(rrcarrb3$PriceSplit),'PriceSplit'] <- 'Low'

pcart <- ggplot(rrcarrb3,
       aes(x=Precip, y = AreaP, colour = PriceSplit)) +
  geom_point() + 
  geom_smooth(method=lm, se=F, size=.5) +
  scale_colour_manual(values=c('#cc4c02','#fe9929'),
                      name='Price Split') +
  ylab('Percent Area Irrigated') + 
  xlab('Precipitation (cm)') + 
  theme_bw() + theme(axis.text=element_text(size=10),
                     legend.text=element_text(size=10),
                     legend.title = element_text(size=10),
                     axis.title=element_text(size=11),
                     legend.position=c(0.2,.8),
                    panel.grid.major = element_blank(),
             panel.grid.minor = element_blank())
pcart

# run regressions 
lms <- rrcarrb3 %>% 
  group_by(PriceSplit) %>% 
  do(model = lm(Area ~ Precip, data = .))

# and model R2 values
rsq <- lms %>% glance(model)
rsq
```

## Combo Figs
grid layout ref http://www.sthda.com/english/wiki/ggplot2-easy-way-to-mix-multiple-graphs-on-the-same-page-r-software-and-data-visualization

```{r precipTrendsCombo_rrbrrca, fig.height=3, fig.width = 6.5, dpi=600, dev=c('png','pdf')}
grid.newpage()
grid.draw(cbind(ggplotGrob(depth), ggplotGrob(pcart), size = "last"))
```

```{r fig3Combo_noGrid_rrbrrca, fig.height=6.5, fig.width = 6, dpi=600, dev=c('png','pdf')}
# grid.newpage()
# # Create layout : nrow = 2, ncol = 2
# pushViewport(viewport(layout = grid.layout(2,2)))
#              
# # A helper function to define a region on the layout
# define_region <- function(row, col){
#   viewport(layout.pos.row = row, layout.pos.col = col)
# } 
# # Arrange the plots
# print(ptime, vp=define_region(1, 1:2))
# print(depth, vp = define_region(2, 1))
# print(pcart, vp = define_region(2, 2))  

grid.arrange(ptime2, depth, pcart, layout_matrix = rbind(c(1,1), c(2,3)),
             heights=c(1.45,1))
```



