---
title: "Confusion Tables"
author: "Jill Deines"
date: "Sunday, November 20, 2016"
output: 
  html_document:
    toc: yes
---

Goal: Accuracy assessment via confusion tables

```{r knitrOpts, echo=FALSE}
library(knitr)
opts_chunk$set(cache=FALSE, cache.path='cache/5.0_confusion/',
               fig.path='figure/5.0_confusion/')
```

**R Packages Needed**

```{r packages, message=FALSE, echo=TRUE}
library(rgdal)
library(caret)


# source project helper functions
source('functions/GeeCsvToSpdf.R')
source('functions/prettyConfusion.R')
```

## 2001
From Brian Wardlow's Kansas dataset of validation points, clipped to the RRB and manually cleaned for bad location points.

```{r Wardlow}
# load data
dataDir <- 'C:/Users/deinesji/Google Drive/GEE_tableExports/Accuracy_PointsforConfusionTables_test1_clean'
file2001 <- '2001_Wardlow_confusionData_RRB_test2_nosoy_noag.csv'
wardlow01 <- read.csv(paste(dataDir, file2001, sep='/'))

# adjust categories for a binary evaluation (irrigated/not irrigated)
wardlow01$binaryClassification <- wardlow01$classification
wardlow01$binaryClassification[wardlow01$binaryClassification == 2] <- 0 

wardlow01$binaryClassCats <- wardlow01$ClassCats
wardlow01$binaryClassCats[wardlow01$binaryClassCats == 2] <- 0 

# lookup table for classification codes
predictedCode = data.frame(classification = c(0:2),
                       predictedClass = c('Dryland','Irrigated','Noncrop'))
referenceCode = data.frame(ClassCats = c(0:2),
                       referenceClass = c('Dryland','Irrigated','Noncrop'))

# binary tables
predCodeBinary = data.frame(binaryClassification = c(0:1),
                       predictedClassBinary = c('NotIrrigated','Irrigated'))
refCodeBinary = data.frame(binaryClassCats = c(0:1),
                       referenceClassBinary = c('NotIrrigated','Irrigated'))

# add category names to df via merge
wardlow01 <- merge(wardlow01, predictedCode)
wardlow01 <- merge(wardlow01, referenceCode)
wardlow01 <- merge(wardlow01, predCodeBinary)
wardlow01 <- merge(wardlow01, refCodeBinary)

# confusion table all 3 classes
wardlowConfusion <- confusionMatrix(data = wardlow01$predictedClass, 
                                      reference = wardlow01$referenceClass)

# confusion table binary
wardlowConfusionBin <- confusionMatrix(data = wardlow01$predictedClassBinary, 
                                      reference = wardlow01$referenceClassBinary)

# make it pretty : all 3 classes
wardlow.ct <- prettyConfusion(wardlowConfusion)
kable(wardlow.ct, digits=1)

# make it pretty : binary
wardlow.ct.b <- prettyConfusionBinary(wardlowConfusionBin)
kable(wardlow.ct.b, digits=1)
```

## 2007
Jill's points for the Middle Republican 2007; I think I used 60% of the points for "validation" and am saving 40% for "test"

```{r midRep2007}
# load data
dataDir <- 'C:/Users/deinesji/Google Drive/GEE_tableExports/Accuracy_PointsforConfusionTables_test1_clean'
file2007 <- '2007_MidRep60_confusionData_RRB_test2_nosoy_noag.csv'
midRep07 <- read.csv(paste(dataDir, file2007, sep='/'))

# adjust categories for a binary evaluation (irrigated/not irrigated)
midRep07$binaryClassification <- midRep07$classification
midRep07$binaryClassification[midRep07$binaryClassification == 2] <- 0 

midRep07$binaryClassNum <- midRep07$classNum
midRep07$binaryClassNum[midRep07$binaryClassNum == 2] <- 0 

# lookup table for classification codes
predictedCode = data.frame(classification = c(0:2),
                       predictedClass = c('Dryland','Irrigated','Noncrop'))
referenceCode = data.frame(classNum = c(0:2),
                       referenceClass = c('Dryland','Irrigated','Noncrop'))

# binary tables
predCodeBinary = data.frame(binaryClassification = c(0:1),
                       predictedClassBinary = c('NotIrrigated','Irrigated'))
refCodeBinary = data.frame(binaryClassNum = c(0:1),
                       referenceClassBinary = c('NotIrrigated','Irrigated'))

# add category names
midRep07 <- merge(midRep07, predictedCode)
midRep07 <- merge(midRep07, referenceCode)
midRep07 <- merge(midRep07, predCodeBinary)
midRep07 <- merge(midRep07, refCodeBinary)

# confusion table
midRep07confusion <- confusionMatrix(data = midRep07$predictedClass, 
                                      reference = midRep07$referenceClass)

# binary
midRep07confusionBinary <- confusionMatrix(data = midRep07$predictedClassBinary, 
                                      reference = midRep07$referenceClassBinary)

# make it pretty
midrep07.ct <- prettyConfusion(midRep07confusion)
kable(midrep07.ct, digits=1)

# binary one
midrep07.ct.b <- prettyConfusionBinary(midRep07confusionBinary)
kable(midrep07.ct.b, digits=1)
```


## 2010
Jill's points for the Middle Republican 2010; I think I used 20% of the points for "validation" and the other 80% were training


```{r midRep2010}
# load data
dataDir <- 'C:/Users/deinesji/Google Drive/GEE_tableExports/Accuracy_PointsforConfusionTables_test1_clean'
file2010 <- '2010_MidRep60_confusionData_RRB_test2_nosoy_noag.csv'
midRep10 <- read.csv(paste(dataDir, file2010, sep='/'))

# adjust categories for a binary evaluation (irrigated/not irrigated)
midRep10$binaryClassification <- midRep10$classification
midRep10$binaryClassification[midRep10$binaryClassification == 2] <- 0 

midRep10$binaryClassNum <- midRep10$classNum
midRep10$binaryClassNum[midRep10$binaryClassNum == 2] <- 0 


# lookup table for classification codes
predictedCode = data.frame(classification = c(0:2),
                       predictedClass = c('Dryland','Irrigated','Noncrop'))
referenceCode = data.frame(classNum = c(0:2),
                       referenceClass = c('Dryland','Irrigated','Noncrop'))

# binary tables
predCodeBinary = data.frame(binaryClassification = c(0:1),
                       predictedClassBinary = c('NotIrrigated','Irrigated'))
refCodeBinary = data.frame(binaryClassNum = c(0:1),
                       referenceClassBinary = c('NotIrrigated','Irrigated'))

# add category names
midRep10 <- merge(midRep10, predictedCode)
midRep10 <- merge(midRep10, referenceCode)
midRep10 <- merge(midRep10, predCodeBinary)
midRep10 <- merge(midRep10, refCodeBinary)

# confusion table
midRep10confusion <- confusionMatrix(data = midRep10$predictedClass, 
                                      reference = midRep10$referenceClass)

# confusion table
midRep10confusionBinary <- confusionMatrix(data = midRep10$predictedClassBinary, 
                                      reference = midRep10$referenceClassBinary)

# make it pretty
midrep10.ct <- prettyConfusion(midRep10confusion)
kable(midrep10.ct, digits=1)

midrep10.ct.b <- prettyConfusionBinary(midRep10confusionBinary)
kable(midrep10.ct.b, digits=1)
```
