---
title: "Statistical Model"
author: "Jill Deines"
date: "February 27, 2017"
output: 
  html_document:
    toc: yes
---

Goal: Investigate relationship among irrigated area, precip, pumping volume, and crop prices over time 

update 4/9/2017: added total study area irrigated area panel to time series plot


```{r knitrOpts, echo=FALSE}
library(knitr)
opts_chunk$set(cache=FALSE, 
               fig.path='figure/6.05_statModels_test4/')
```

**R Packages Needed**

```{r packages, message=FALSE, echo=TRUE}
library(tidyverse) # for ggplot2, tidyr, dplyr
library(broom)
library(GGally)
library(corrplot)
library(Hmisc)
library(party) #carts
source('functions/stat_smooth_func_all.R')
library(RColorBrewer)
```

## User Vars
Derived dataset source files

* total irrigated area, full region: written out in NassYearTrends chunk of 6.00_IrrigationTrends
* corn prices: produced in first chunks of this script
* RRCA domain irrigation and precip in 4.21_Figures_volumeVsArea_randFor... in the 'loadAreaData' chunk
* RRCA pumping volume: processed in 4.21_Figures_volumeVsArea_randFor... 'loadPumpingData' chunk


```{r userVars}
# data directory
derivedDir <- 'C:/Users/deinesji/Dropbox/1PhdJill/hpa/irrigation/data/derivedDatasets'

# RRCA domain irrigation and precipitation by year
Irr.Ppt.file <- 'RRCA_fullModel_area-rfTest4_ppt-dec-aug.csv'

# full study area domain irrigation
Irr.fullExtent.file <- 'FullStudyExtent_irrigation_test4.csv'

# corn prices
cornFile <- 'price_nass_corn_1996-2016.csv'

# RRCA volume
RRCA.volume.file <- 'RRCA_fullModel_irrigation_volume.csv'
```

## Data wrangling
(jump to Analyses/loaddata chunk after first time)

### NASS crop price data

First, make a table of inflation adjustments relative to 2015, the final year available in my NASS price dataset. Write out so this is static and not dependent on external download

```{r inflationTable, eval=FALSE}
# adjust for inflation
library(lubridate) 
library(dplyr) 

# download the consumer price index and get adjustments 
monthly_cpi <- read.csv("http://research.stlouisfed.org/fred2/data/CPIAUCSL.csv",
                        header = TRUE)
monthly_cpi$cpi_year <- year(monthly_cpi$DATE)
yearly_cpi <- monthly_cpi %>% group_by(cpi_year) %>% summarize(cpi = mean(VALUE))
yearly_cpi$adj_factor_2015 <- yearly_cpi$cpi/yearly_cpi$cpi[yearly_cpi$cpi_year == 2015]
# for posterity
yearly_cpi$adj_factor_2016 <- yearly_cpi$cpi/yearly_cpi$cpi[yearly_cpi$cpi_year == 2016]
yearly_cpi$adj_factor_2017 <- yearly_cpi$cpi/yearly_cpi$cpi[yearly_cpi$cpi_year == 2017]

# write out
nassdir <- 'C:/Users/deinesji/Dropbox/1PhdJill/hpa/irrigation/data/nass/price'
write.csv(yearly_cpi, paste0(nassdir, '/cpi_priceAdjustments.csv'), 
          row.names=F)
```

Price data was downloaded manually from the NASS quickstats tool (https://quickstats.nass.usda.gov/) for the following query: SURVEY - CROPS - FIELD CROPS - CORN/SOYBEANS/WHEAT/HAY/SORGHUM - PRICE RECEIVED - 

* CORN, GRAIN - PRICE RECEIVED, MEASURED IN $ / BU
* HAY, PRICE RECEIVED, MEASURED IN $ / TON (short ton)
* SORGHUM, GRAIN - PRICE RECEIVED, MEASURED IN $ / CWT (hundredweight)
* SOYBEANS - PRICE RECEIVED, MEASURED IN $ / BU
* WHEAT - PRICE RECEIVED, MEASUTED IN $ / BU

- TOTAL - NATIONAL - US TOTAL - 1996:2016 - ANNUAL - YEAR


Here, I load the data and clean it to have consistent pricing across crops and years

```{r nassPrice, eval=FALSE}
# load nass
nassdir <- 'C:/Users/deinesji/Dropbox/1PhdJill/hpa/irrigation/data/nass/price'
price0 <- read.csv(paste0(nassdir, '/nass_prices_5crops_national_1996-2016.csv'))

# load consumer price index adjustments
cpi <- read.csv(paste0(nassdir, '/cpi_priceAdjustments.csv'))

# make a units columns
unitkey <- data.frame(Commodity = c('CORN','HAY','SORGHUM','SOYBEANS','WHEAT'),
                      unit = c('BU','TON','CWT','BU','BU'))
price0 <- merge(price0,unitkey)

# remove silly columns
price0 <- price0[,c('Commodity','Year','Value','unit')]

# add the adjustment factor
price <- merge(price0, cpi[,c('cpi_year','adj_factor_2015','adj_factor_2016')], 
               by.x = 'Year', by.y = 'cpi_year')

# adjust prices (note: matched trials at https://data.bls.gov/cgi-bin/cpicalc.pl)
price$price_2015 <- price$Value / price$adj_factor_2015
price$price_2016 <- price$Value / price$adj_factor_2016

# plot it - unadjusted
ggplot(data = price, aes(x=Year, y = Value)) +
  geom_line() +
  facet_wrap(~Commodity, scales='free_y') + theme_bw() +
  ggtitle('Nomial (unadjusted) Crop Prices') 


# plot it
ggplot(data = price, aes(x=Year, y = price_2016)) +
  geom_line() +
  facet_wrap(~Commodity, scales='free_y') + theme_bw() +
  ggtitle('Inflation-adjusted Crop Prices')

# extract only corn
cornprice <- price[price$Commodity == 'CORN',]

ggplot(data = cornprice, aes(x=Year, y = price_2016)) +
  geom_line() + theme_bw() +
  ggtitle('Inflation-adjusted Corn Prices')

# write out dataset
write.csv(cornprice, paste0(derivedDir, '/price_nass_corn_1996-2016.csv'), 
          row.names=F)
```

## Analyses

### Load Data
loads datasets produced/cleaned from other scripts

```{r loadData}
# usda nass annual corn price, $ / bushel
cornprice <- read.csv(paste0(derivedDir, '/',cornFile))

# irrigation area and precipitation
area.state <- read.csv(paste0(derivedDir,'/',Irr.Ppt.file))
# aggregate area across states - sum
area <- aggregate(irrigated_km2 ~ year, data = area.state, FUN = 'sum')
# aggregate ppt across states - mean
ppt <- aggregate(pr_ann ~ year, data = area.state, FUN = 'mean')
# merge back
areappt <- merge(area, ppt)

# irrigation volume ----------
volume.state <- read.csv(paste0(derivedDir, '/',RRCA.volume.file))
# aggregate volume across states - sum
volume <- aggregate(total_km3 ~ year, data = volume.state, FUN = 'sum', na.rm=T)

# merge all ---------------
# merge price
area.ppt.price <- merge(areappt, cornprice[,c('Year','price_2016')], 
                        by.x = 'year', by.y = 'Year')

# merge volume
rrca <- merge(area.ppt.price,volume, all=T)
# rename columns
names(rrca) <- c('Year','Area_km2','Precip_mm','Price','Volume_km3')

# add a price lag
rrca$priceLag1 <- cornprice$price_2016[3:20]

# add total regional area
area.full <- read.csv(paste0(derivedDir,'/',Irr.fullExtent.file))
rrca$AArea_km2_full <- area.full$area_km2
```

### Exploratory analyses: RRCA domain

Average application depth: Divide irrigated pumping by area (average application depth); plot vs precip expressed in mm

what proportion of water demand they are satisfying?

```{r applicationDepth, fig.height=3, fig.width=3}
# calculate depth of water applied
rrca$Depth_m <- rrca$Volume_km3/rrca$Area_km2 * 1000

ggplot(rrca, aes(Precip_mm, Depth_m)) +
  geom_point() + theme_bw() +
  stat_smooth(method = 'lm', se=F) +
  ylab('Depth (m)') + xlab('Precip (mm)') +
  ggtitle('Precip vs Application Depth')

depthlm <- lm(Depth_m ~ Precip_mm, data = rrca)
summary(depthlm)
```

And plot all time series

```{r timeSeriesPlot, fig.height = 6, fig.width = 6}
# data frame for plotting
rrcaPlot <- rrca

# items to plot
plotUs <- c('AArea_km2_full','Area_km2','Depth_m','Precip_mm','Price')

# gather
plotdata <- gather_(rrcaPlot, key_col = 'variable', value_col = 'value',
                          gather_cols = plotUs)

# plots with no facet labels
ggplot(data = plotdata,
      aes(x = Year, y = value, colour=variable)) +
  geom_line() + geom_point() +
  scale_colour_manual(values=c('black','black','red','blue','orange')) + 
  facet_wrap(~variable, nrow=5, scales='free_y') +
  scale_x_continuous(name = '', 
                  breaks = c(2000,2002,2004,2006,2008,2010,2012,2014,2016), 
                 labels = c(2000,2002,2004,2006,2008,2010,2012,2014,2016)) +
  ylab('') +
  theme_bw() + theme(legend.position = 'none',
                     strip.background=element_blank(), # comment out to add labels
                     strip.text.x = element_blank(),   # comment out to add labels
                     text = element_text(size = 15))

# plots with trendlines
ggplot(data = plotdata,
      aes(x = Year, y = value, colour=variable)) +
  geom_line() + geom_point() +
  scale_colour_manual(values=c('black','black','red','blue','orange')) + 
  facet_wrap(~variable, nrow=5, scales='free_y') +
  stat_smooth_func(geom="text",method="lm",hjust=0,parse=TRUE, 
                 xpos = 2004) +
  stat_smooth(method = 'lm', se=F) +
  scale_x_continuous(name = '', 
                  breaks = c(2000,2002,2004,2006,2008,2010,2012,2014,2016), 
                 labels = c(2000,2002,2004,2006,2008,2010,2012,2014,2016)) +
  ylab('') +
  theme_bw() + theme(legend.position = 'none',
                     strip.background=element_blank(), # comment to add labels
                     strip.text.x = element_blank(),   # comment to add labels
                     text = element_text(size = 15))

# plots with trendlines no text
ggplot(data = plotdata,
      aes(x = Year, y = value, colour=variable)) +
  geom_line() + geom_point() +
  scale_colour_manual(values=c('black','black','red','blue','orange')) + 
  facet_wrap(~variable, nrow=5, scales='free_y') +
  stat_smooth(method = 'lm', se=F) +
  scale_x_continuous(name = '', 
                  breaks = c(2000,2002,2004,2006,2008,2010,2012,2014,2016), 
                 labels = c(2000,2002,2004,2006,2008,2010,2012,2014,2016)) +
  ylab('') +
  theme_bw() + theme(legend.position = 'none',
                     strip.background=element_blank(), # comment to add labels
                     strip.text.x = element_blank(),   # comment to add labels
                     text = element_text(size = 15))

```

Run a quick set of individual regressions on these time series

```{r timeSeriesLMs}
# run regressions by county
lms <- plotdata %>% 
  group_by(variable) %>% 
  do(model = lm(value ~ Year, data = .))

# tidy up with broom: model coefficients
coeffs <- lms %>% tidy(model)
coeffs <- coeffs[coeffs$term == 'Year',]
coeffs

# and model R2 values
rsq <- lms %>% glance(model)
rsq
```



Let's look at some preliminary correlations for the RRCA domain. 

```{r corrPlots, fig.height = 5, fig.width=5}
# remove the full domain data
rrca <- rrca[,-which(names(rrca) %in% c('AArea_km2_full'))]

# cor plots
varcor <- cor(rrca, use='complete.obs')
corrplot(varcor, method='color', addCoef.col='black', tl.col='black',
         type = 'lower',
         col = colorRampPalette(brewer.pal(7,'RdBu'))(200))


#ggcorr(rrca, palette = "RdBu", label = TRUE, digits=3)

# sig levels
corsig <- rcorr(as.matrix(rrca[complete.cases(rrca),]), type = 'pearson')
options(scipen = 999)
corsig$P
```

Interpretation:

* Precip has inverse correlations with irrigation volume (significant) and irrigation area (non-significant)
* Volume and area are negatively correlated (p = 0.077)
* Price Lag 1 correlated with area
* Depth negatively corrected with area, precip; positive with volume

```{r scatterplots, fig.height=5.5, fig.width=8.5}
# an extra function to add trendlines
my_fn <- function(data, mapping, ...){
  p <- ggplot(data = data, mapping = mapping) + 
    geom_point() + 
    geom_smooth(method=loess, fill="red", color="red", ...) +
    geom_smooth(method=lm, fill="blue", color="blue", ...)
  p
}

# scatterplot matrix
ggpairs(rrca, lower=list(continuous = my_fn)) 
```

Hm. A two step model? Predict application depth from precip; predict irrigated area from depth and price? Or a hierarchical model where depth comes from precip...


### Auto-correlation


```{r autocorrelation, fig.height = 5.5, fig.width=8, eval=FALSE}
# test for autocorrelation in time series
vars <- names(rrca)

# test all vars
par(mfrow=c(3,3))
acfList <- list()
#names(acfList) <- vars
for (var in vars){
  acf(rrca[,var], main = var, na.action = na.pass)
}
```

Area and price are autocorrelated. Calculate first differences instead.

```{r autocor2, fig.height = 5.5, fig.width = 6, eval=FALSE}
# plot first differences: calculate
rrca.firstdiff <- as.data.frame(matrix(, nrow = nrow(rrca)-1, ncol=length(vars)))
names(rrca.firstdiff) <- vars
for (var in vars) {
 rrca.firstdiff[,var] <- diff(ts(rrca[,var]), differences=1)
}

rrca.firstdiff$year <- 2000:2016
rrca.firstdiff <- rrca.firstdiff[ , -which(names(rrca.firstdiff) %in% c('Price'))]

# longify

firstdiff <- gather_(rrca.firstdiff, key_col = 'variable', value_col = 'value',
                     plotUs)

ggplot(data = firstdiff,
      aes(x = year, y = value, colour=variable)) +
  geom_line() + geom_point() +
  scale_colour_manual(values=c('black','red','blue','orange')) + 
  facet_wrap(~variable, nrow=5, scales='free_y') +
  scale_x_continuous(name = '', 
                     breaks = c(2000,2002,2004,2006,2008,2010,2012,2014,2016), 
                     labels = c(2000,2002,2004,2006,2008,2010,2012,2014,2016)) +
  ylab('') + ggtitle('first difference') +
  theme_bw() + theme(legend.position = 'none',
                     strip.background=element_blank(), # comment to add labels
                     strip.text.x = element_blank(),   # comment to add labels
                     text = element_text(size = 16))
```


### CaRT
I picked 'ctree' function the party package because of the nice-ish output on an R Blogger's post, but Andy subsequently suggested it as well. Look at that!

From 'party' package

```{r cart, fig.height = 5.5, fig.width = 7.5}
rrca$farmBill <- c(rep(1,3), # 1999-2001
                   rep(2,6), # 2002-2007
                   rep(3,6), # 2008-2013
                   rep(4,3)) # 2014-2016

rrcaPlot$farmBill <- c(rep(1,3), # 1999-2001
                   rep(2,6), # 2002-2007
                   rep(3,6), # 2008-2013
                   rep(4,3)) # 2014-2016

# new cart, whole study area ---------------------------------------------
 # control tree
controls <- ctree_control(minsplit = 2, minbucket = 3, mincriterion = .5)

# run tree
ct <- ctree(AArea_km2_full ~ Precip_mm + priceLag1, data = rrcaPlot, 
            controls = controls)
plot(ct, main="Conditional Inference Tree")

# new cart, RRCA area ---------------------------------------------
 # control tree
controls <- ctree_control(minsplit = 2, minbucket = 3, mincriterion = .5)

# run tree
ct <- ctree(Area_km2 ~ Precip_mm + priceLag1, data = rrca, 
            controls = controls)
plot(ct, main="Conditional Inference Tree")



## cart without 2016  ------------------------
 # control tree
controls <- ctree_control(minsplit = 2, minbucket = 2, mincriterion = .5)

# run tree
ct <- ctree(Area_km2 ~ Precip_mm + priceLag1, data = rrca[1:17,], 
            controls = controls)
plot(ct, main="Conditional Inference Tree")


# with the farm bill ---------------------------------------------
 # control tree
controls <- ctree_control(minsplit = 2, minbucket = 2, mincriterion = .6)

# run tree
ct <- ctree(Area_km2 ~ Precip_mm + priceLag1 + farmBill, data = rrcaPlot, 
            controls = controls)
plot(ct, main="Conditional Inference Tree")

```


